---
title: "Metabolomics_Loreal"
output: html_document
date: "2024-09-25"
author: "Britta De Pessemier, adapted by Yang Chen 2025-0108"
---

```{r packages, message=FALSE, warning=FALSE}
# Load Required Libraries
library(tidyverse)
library(dplyr)
library(stringr)
library(tibble)

# Statistical Analysis
library(vegan)
library(rstatix)
library(lmerTest)
library(caret)

# Visualization
library(ggplot2)
library(ggpubr)
library(cowplot)
library(RColorBrewer)
library(svglite)

# Specialized Tools
library(mixOmics)
library(UpSetR)
library(homologueDiscoverer) # devtools::install_github("kevinmildau/homologueDiscoverer")
library(MassTools)           # devtools::install_github("mjhelf/MassTools")
```

```{r}
# Import needed files from metabolomics run 3
feature_table <- read_csv("../Data/metabolomics/Run3_10252024/FBMN_mgf_Lami_10252024_quant.csv")

metadata <- read_csv("../Metadata/metadata_Tyler_reordered.csv")

sample_order <- read.csv("../Data/metabolomics/Run3_10252024/Sequence_order_10262024.csv")

# With suspects library
annotations <- read.delim("../Data/metabolomics/Run3_10252024/merged_results_with_gnps.tsv")

# Without suspects library
#annotations <- read.delim("../Data/metabolomics/Run3_10252024/merged_results_with_gnps_withoutsuspects.tsv")
annotations$X.Scan. <- as.character(annotations$X.Scan.)
canopus <- read_tsv("../Data/metabolomics/Run3_10252024/canopus_structure_summary.tsv")
canopus$featureId <- as.character(canopus$featureId)

mapping_file <- read.csv("../Data/metabolomics/Run3_10252024/Mapping_10262024.csv")
```

## Process and integrate metabolomics data by cleaning, merging, and aligning feature tables, metadata, annotations, and sample order files while identifying and handling duplicates, missing values, and sample exclusions for downstream analysis

```{r}
# Find duplicate sample names in the metadata
# Create a subset of rows in the metadata where sample_name appears more than once
duplicates_in_metadata <- metadata[metadata$sample_name %in% metadata$sample_name[duplicated(metadata$sample_name)], ]

# Replace underscores with periods in the mapping_file$Manifest_ID_Sample_Name column
mapping_file$Manifest_ID_Sample_Name <- gsub("_", ".", mapping_file$Manifest_ID_Sample_Name)

# Merge metadata with mapping_file based on sample_name
# Ensure sample_name in metadata matches Manifest_ID_Sample_Name in mapping_file
metadata_2 <- metadata %>% left_join(mapping_file, by = c("sample_name" = "Manifest_ID_Sample_Name"))

# Add a new column SampleID to metadata_2 using Filename_ID
metadata_2$SampleID <- metadata_2$Filename_ID

# Identify duplicate sample names in the updated metadata_2
duplicates_in_metadata <- metadata_2[metadata_2$sample_name %in% metadata_2$sample_name[duplicated(metadata_2$sample_name)], ]

# Reorder columns in metadata_2 to place SampleID as the first column
metadata_2 <- metadata_2[, c("SampleID", setdiff(names(metadata_2), "SampleID"))]
```

```{r}
# Identify rows in mapping_file where Filename_ID is not present in metadata_2$SampleID
missing_sample_ids <- mapping_file[!(mapping_file$Filename_ID %in% metadata_2$SampleID), ]

# Create a new data frame for missing SampleIDs
missing_sample_ids_df <- data.frame(SampleID = missing_sample_ids$Filename_ID)

# Add missing SampleIDs to metadata_final with other columns filled with NA
missing_rows <- data.frame(SampleID = missing_sample_ids_df$SampleID, 
                           matrix(NA, nrow = nrow(missing_sample_ids_df), ncol = ncol(metadata_2) - 1))
colnames(missing_rows) <- colnames(metadata_2)

# Remove any duplicate rows that might already exist in metadata_2
missing_rows <- missing_rows[!(missing_rows$SampleID %in% metadata_2$SampleID), ]

# Append missing rows to metadata_2 to create metadata_final
metadata_final <- rbind(metadata_2, missing_rows)

# Add a new column Sample_Type to metadata_final based on SampleID patterns
metadata_final <- metadata_final %>%
  mutate(Sample_Type = case_when(
    str_detect(SampleID, regex("wash", ignore_case = TRUE)) ~ "Wash",
    str_detect(SampleID, regex("Blank", ignore_case = TRUE)) ~ "Blank",
    str_detect(SampleID, "^QC_P[0-9]+_[0-9]+$") | str_detect(SampleID, regex("6mix", ignore_case = TRUE)) ~ "6mix",
    str_detect(SampleID, regex("Pool", ignore_case = TRUE)) ~ "Pooled samples",
    str_detect(SampleID, regex("CTL", ignore_case = TRUE)) ~ "CTL",
    TRUE ~ "samples"  # Default to 'samples' for SampleIDs that don't match the others
  ))

# Print or inspect the updated metadata_final with Sample_Type column
head(metadata_final)
```

```{r}
# Summarize counts of each Sample_Type in metadata_final
cat("Number of 'samples':", sum(metadata_final$Sample_Type == "samples"), "\n")
cat("Number of '6mix':", sum(metadata_final$Sample_Type == "6mix"), "\n")
cat("Number of 'Pooled samples':", sum(metadata_final$Sample_Type == "Pooled samples"), "\n")
cat("Number of 'CTL':", sum(metadata_final$Sample_Type == "CTL"), "\n")
cat("Number of 'Blank':", sum(metadata_final$Sample_Type == "Blank"), "\n")

# Summarize counts of each category_label in metadata
cat("Number of 'Healthy':", sum(metadata$category_label == "Healthy"), "\n")
cat("Number of 'Acne Lesional':", sum(metadata$category_label == "Acne Lesional"), "\n")
cat("Number of 'Acne Non-lesional':", sum(metadata$category_label == "Acne Non-lesional"), "\n")
```

```{r}
# L'Oreal Data Integration
# Merge metadata_final with sample_order by matching SampleID and File.Name
metadata_final <- metadata_final %>% left_join(sample_order, by = c("SampleID" = "File.Name"))

# Identify unmatched SampleIDs between metadata_final and sample_order
unmatched_metadata <- metadata_final %>%
  anti_join(sample_order, by = c("SampleID" = "File.Name"))

# View the unmatched metadata
unmatched_metadata

unmatched_sample_order <- sample_order %>%
  anti_join(metadata_final, by = c("File.Name" = "SampleID"))

# View the result
unmatched_sample_order
```

```{r}
# Check for duplicated SampleIDs after the join
duplicated_samples <- metadata_final %>%
  group_by(SampleID) %>%
  filter(n() > 1)

# Display the duplicated SampleIDs
cat("Duplicated SampleIDs after the join:\n")
print(duplicated_samples)

# Optionally, count the number of duplicates
num_duplicates <- nrow(duplicated_samples)
cat("\nNumber of duplicated SampleIDs after the join: ", num_duplicates, "\n")
```

```{r}
# Identify unmatched SampleIDs between sample_order and metadata_final
# Create a subset of rows from metadata_final that are not in sample_order
unmatched_metadata <- metadata_final %>%
  anti_join(sample_order, by = c("SampleID" = "File.Name"))

# View the unmatched metadata
unmatched_metadata

# Extract and prepare feature information from the feature_table
info_feature <- feature_table %>% dplyr::select(1:3,7)
colnames(info_feature) <- c("Feature", "mz", "RT", "Corr_ID")
info_feature$Feature <- as.character(info_feature$Feature)

# Merge feature information with annotations and CANOPUS data
info_feature_complete <- info_feature %>% 
  left_join(annotations, by = c("Feature" = "X.Scan.")) %>%
  dplyr::select(1:4,18,24) %>% 
  left_join(canopus %>% distinct(featureId, .keep_all = TRUE), by = c("Feature" = "featureId"))
```

```{r}
# Extract data table from feature_table
# Transform and clean up the data for analysis
data <- feature_table %>% 
  column_to_rownames("row ID") %>% dplyr::select(contains("Peak")) %>% 
  t() %>% as.data.frame() %>% rownames_to_column("SampleID") %>% 
  arrange(SampleID) %>% distinct(SampleID, .keep_all = TRUE)

# Clean up SampleID to remove unnecessary text
data$SampleID <- gsub(".mzML Peak area", "", data$SampleID)
```

```{r}
# Find the rows in mapping_file where Filename_ID is not present in metadata_final$SampleID
missing_sample_ids <- data[!(data$SampleID %in% metadata_final$SampleID), ]

# Find the rows in mapping_file where Filename_ID is not present in metadata_final$SampleID
missing_samples <- metadata_final[!(metadata_final$SampleID %in% data$SampleID), ]
# LAMI.RD318.D4.C1 did not run

metadata_ids <- metadata_final$SampleID
data_ids <- data$SampleID

# 1. Find SampleIDs in metadata_final but not in data
missing_in_data <- setdiff(metadata_ids, data_ids)

# 2. Find SampleIDs in data but not in metadata_final
missing_in_metadata <- setdiff(data_ids, metadata_ids)

# 3. Check for duplicated SampleIDs in metadata_final
duplicates_in_metadata <- metadata_ids[duplicated(metadata_ids)]

# 4. Check for duplicated SampleIDs in data
duplicates_in_data <- data_ids[duplicated(data_ids)]

# Print results
cat("SampleIDs in metadata_final but not in data:\n")
print(missing_in_data)

cat("\nSampleIDs in data but not in metadata_final:\n")
print(missing_in_metadata) # Should be none

cat("\nDuplicated SampleIDs in metadata_final:\n")
print(duplicates_in_metadata) # Should be none

cat("\nDuplicated SampleIDs in data:\n")
print(duplicates_in_data) # Should be none
```

```{r}
# Clean up SampleID to remove unnecessary text
data$SampleID <- gsub(".mzML Peak area", "", data$SampleID)

# Filter out specific samples from the dataset (blanks, QC mixes)
data_final <- data %>% 
  dplyr::filter(!(SampleID %in% c("blank_P2_0A_prerun", "blank_P3_0A_prerun_20241022011201", "blank_P3_0B_prerun", "blank_P1_0A_prerun", "QC_P1_6mix_0", "QC_P1_Pool_0", "QC_P1_6mix_0_20241019144629", "blank_P1_0A_prerun_20241019143524")))

# Check for missing samples in the dataset
samples_to_remove <- c("blank_P2_0A_prerun", "blank_P3_0A_prerun_20241022011201", "blank_P3_0B_prerun",
                       "blank_P1_0A_prerun", "QC_P1_6mix_0", "QC_P1_Pool_0","QC_P1_6mix_0_20241019144629", "blank_P1_0A_prerun_20241019143524")

# Identify samples that are in the list but not in the dataset
missing_samples <- setdiff(samples_to_remove, data$SampleID)

# Issue a warning if there are missing samples
if (length(missing_samples) > 0) {
  warning(paste("The following samples are not present in the dataset:", paste(missing_samples, collapse = ", ")))
}
```





### QC analysis and plots

```{r}
# Calculate Total Ion Count (TIC)
data_TIC <- data.frame(TIC = rowSums(data_final %>% column_to_rownames("SampleID"))) %>% 
  rownames_to_column("SampleID") %>% left_join(sample_order, by = c("SampleID" = "File.Name"))

#Plots a scatter plot of the Order vs. TIC
data_TIC %>% dplyr::filter(!(str_detect(SampleID, regex("Blank|6mix|Pool|wash|CTL|QC", ignore_case = TRUE)))) %>%
  left_join(metadata_final) %>%
  ggscatter("Order", "TIC", color = "host_subject_id", add = "reg.line") +
  stat_cor()

ggsave("../Figures/metabolomics_Figures/qc/TICvsOrder.png", plot = last_plot(), dpi = 600, width = 8, height = 6, units = "in")
```

```{r}
# Check sample types based on Total Ion Count (TIC)
sample_tic <- data_TIC %>% 
  dplyr::filter(!(str_detect(SampleID, regex("Blank|6mix|Pool|wash|CTL|QC", ignore_case = TRUE)))) %>% 
  summarise(mean(TIC))

pool_tic <- data_TIC %>% 
  dplyr::filter(str_detect(SampleID, regex("Pool", ignore_case = TRUE))) %>% 
  summarise(mean(TIC))

six_tic <- data_TIC %>% 
  dplyr::filter(str_detect(SampleID, "^QC_P[0-9]+_[0-9]+$") | str_detect(SampleID, regex("6mix", ignore_case = TRUE))) %>% 
  summarise(mean(TIC))

blank_tic <- data_TIC %>% 
  dplyr::filter(str_detect(SampleID, regex("blank", ignore_case = TRUE))) %>% 
  summarise(mean(TIC))

ctl_tic <- data_TIC %>% 
  dplyr::filter(str_detect(SampleID, regex("CTL", ignore_case = TRUE))) %>% 
  summarise(mean(TIC))

# Calculate TIC ratios
ratio_tic_pb <- pool_tic / sample_tic
ratio_tic_sb <- sample_tic / blank_tic
ratio_tic_6p <- pool_tic / six_tic

# Visualize TIC over time for different sample types
a0 <- data_TIC %>% dplyr::filter(!(str_detect(SampleID, regex("Blank|6mix|Pool|wash|CTL|QC", ignore_case = TRUE)))) %>% 
  ggscatter("Order", "TIC", add = "reg.line") +
  stat_cor() + ylim(0, 7e9) + ggtitle("TIC over Time for Samples")

a <- data_TIC %>% dplyr::filter((str_detect(SampleID, regex("Pool", ignore_case = TRUE)))) %>% 
  ggscatter("Order", "TIC", add = "reg.line") +
  stat_cor() + ylim(0, 7e9) + ggtitle("TIC over Time for Pooled Samples")

b <- data_TIC %>% dplyr::filter(str_detect(SampleID, "^QC_P[0-9]+_[0-9]+$") | str_detect(SampleID, regex("6mix", ignore_case = TRUE))) %>% 
  ggscatter("Order", "TIC", add = "reg.line") +
  stat_cor() + ylim(0, 7e9) + ggtitle("TIC over Time for 6mix Samples")

c <- data_TIC %>% dplyr::filter((str_detect(SampleID, regex("blank", ignore_case = TRUE)))) %>% 
  ggscatter("Order", "TIC", add = "reg.line") +
  stat_cor() + ylim(0, 7e9) + ggtitle("TIC over Time for Blank Samples")

d <- data_TIC %>% dplyr::filter((str_detect(SampleID, regex("CTL", ignore_case = TRUE)))) %>% 
  ggscatter("Order", "TIC", add = "reg.line") +
  stat_cor() + ylim(0, 7e9) + ggtitle("TIC over Time for Blank Swab Samples (CTL)")

f <- ggarrange(a0, a, b, c, d, ncol = 3, nrow = 2, labels = c("a", "b", "c", "d", "e"))

ggsave("../Figures/metabolomics_Figures/qc/TICvsOrder_all.png", plot = last_plot(), dpi = 600, width = 16, height = 10, units = "in")
```

```{r}
# Identify SampleIDs in data_TIC that are not in data_final
missing_sample_ids <- data_TIC$SampleID[!(data_final$SampleID %in% data_TIC$SampleID)]

# Display or print the missing SampleIDs
print(missing_sample_ids)

# Check for duplicate SampleIDs in data_TIC
duplicates_in_data_TIC <- data_TIC$SampleID[duplicated(data_TIC$SampleID)]

# Display the duplicate SampleIDs
print(duplicates_in_data_TIC)
```

```{r}
# Step 1: Filter by SampleID and count the remaining samples
filtered_by_sampleid <- data_TIC %>%
  dplyr::filter(!(str_detect(SampleID, regex("Blank|6mix|Pool|wash|CTL|QC", ignore_case = TRUE))))
count_after_sampleid <- nrow(filtered_by_sampleid)

# Step 2: Filter by TIC range within the previously filtered dataset and count
data_tic_filtered <- filtered_by_sampleid %>%
  dplyr::filter(TIC >= 6e8 & TIC <= 5e9)
count_after_tic <- nrow(data_tic_filtered)

# Step 3: Identify SampleIDs that do not meet both filtering criteria
# Samples in filtered_by_sampleid but not in data_tic_filtered
excluded_samples <- anti_join(filtered_by_sampleid, data_tic_filtered, by = "SampleID")

# Step 4: Remove these excluded samples from the original data_TIC
data_TIC_updated <- data_TIC %>%
  dplyr::anti_join(excluded_samples, by = "SampleID")

# Print counts for reference
cat("Initial sample count:", nrow(data_TIC), "\n")
cat("Count after SampleID filtering:", count_after_sampleid, "\n")
cat("Final count after TIC filtering:", count_after_tic, "\n")
cat("Remaining samples in data_TIC_final:", nrow(data_TIC_updated), "\n")

# Update data_final by excluding samples listed in excluded_samples
data_final_updated <- data_final %>% 
  dplyr::filter(!(SampleID %in% excluded_samples$SampleID))

# Display filtered data
head(data_TIC_updated)
```


```{r}
# Step 1: Filter out samples that match "Blank", "6mix", "Pool", "wash", "CTL", or "QC" from the subset
# and apply additional filtering to keep only those within the TIC range.
data_tic_filtered <- data_TIC %>%
  dplyr::filter(!(str_detect(SampleID, regex("Blank|6mix|Pool|wash|CTL|QC", ignore_case = TRUE)))) %>%
  dplyr::filter(TIC >= 6e8 & TIC <= 5e9)

# Step 2: Plot the filtered data to visualize TIC over time for the selected samples.
a0 <- data_tic_filtered %>%
  ggscatter("Order", "TIC", add = "reg.line") +
  stat_cor()  +ylim(0, 7e9) +
  ggtitle("TIC over Time for Filtered Samples")

a <- data_TIC %>% dplyr::filter((str_detect(SampleID, regex("Pool", ignore_case = TRUE))))%>% 
  ggscatter("Order", "TIC", add = "reg.line") +
  stat_cor() +ylim(0, 7e9) + ggtitle("TIC over Time for Pooled Samples")

b <- data_TIC %>%   dplyr::filter(str_detect(SampleID, "^QC_P[0-9]+_[0-9]+$") | str_detect(SampleID, regex("6mix", ignore_case = TRUE)))  %>% 
  ggscatter("Order", "TIC", add = "reg.line") +
  stat_cor() +ylim(0, 7e9) + ggtitle("TIC over Time for 6mix Samples")

c <- data_TIC %>% dplyr::filter((str_detect(SampleID, regex("blank", ignore_case = TRUE)))) %>% 
  ggscatter("Order", "TIC", add = "reg.line") +
  stat_cor()  +ylim(0, 7e9) + ggtitle("TIC over Time for Blank Samples")

d <- data_TIC %>%  dplyr::filter((str_detect(SampleID, regex("CTL", ignore_case = TRUE)))) %>% 
  ggscatter("Order", "TIC", add = "reg.line") +
  stat_cor() +ylim(0, 7e9) + ggtitle("TIC over Time for Blank Swab Samples (CTL)")

f <- ggarrange(a0, a, b, c, d, ncol = 3, nrow = 2, labels = c("a", "b", "c", "d", "e"))

# Display the plot
print(a0)

ggsave("../Figures/metabolomics_Figures/qc/TICvsOrder_all_filtered_5e8.png", plot = last_plot(), dpi = 600, width = 16, height = 10, units = "in")
```
```{r}
# Plot a scatter plot of the Order vs. TIC
data_tic_filtered %>% dplyr::filter(!(str_detect(SampleID, regex("Blank|6mix|Pool|wash|CTL|QC", ignore_case = TRUE)))) %>%
  left_join(metadata_final) %>%
  ggscatter("Order", "TIC", color = "host_subject_id", add = "reg.line") +
  stat_cor()

ggsave("../Figures/metabolomics_Figures/qc/TICvsOrder_sampleID.png", plot = last_plot(), dpi = 600, width = 16, height = 10, units = "in")
```

```{r}
# Merge `sample_order` with `mapping_file` to include additional columns (`Filename_ID` and `Plate_Number`) 
# while retaining all rows from `sample_order`.
sample_order_2 <- merge(
  sample_order, 
  mapping_file[, c("Filename_ID", "Plate_Number")], 
  by.x = "File.Name", 
  by.y = "Filename_ID", 
  all.x = TRUE
)

# Merging the `data_TIC` DataFrame with `sample_order_2` based on matching `SampleID` in `data_TIC` with `File.Name` and `Order` with `Ord` in `sample_order_2`
data_tic_plate <- data_TIC %>%
  left_join(sample_order_2, by = c("SampleID" = "File.Name", "Order" = "Order"))

# SampleIDs in data_TIC that did not merge into data_tic_plate (i.e., not present in sample_order_2)
not_in_sample_order <- data_TIC %>%
  anti_join(sample_order_2, by = c("SampleID" = "File.Name", "Order" = "Order"))

# SampleIDs in sample_order_2 that did not merge into data_tic_plate (i.e., not present in data_TIC)
not_in_data_TIC <- sample_order_2 %>%
  anti_join(data_TIC, by = c("File.Name" = "SampleID", "Order" = "Order"))

# Convert Plate_Number to a factor for consistent coloring across all plots
data_tic_plate <- data_tic_plate %>%
  dplyr::mutate(Plate_Number = as.factor(Plate_Number))

# Define the mapping of specific colors to specific Plate_Number values
custom_colors <- c("1" = "#D55E00",  # Plate_Number == 1
                   "2" = "#2B579A",  # Plate_Number == 2
                   "3" = "#A1280B",  # Plate_Number == 3
                   "4" = "#217346")  # Plate_Number == 4

# TIC over Time for Samples (a0)
a0 <- data_tic_plate %>%
  dplyr::mutate(Plate_Number = as.factor(Plate_Number)) %>%
  dplyr::filter(!(str_detect(SampleID, regex("Blank|6mix|Pool|wash|CTL|QC", ignore_case = TRUE)))) %>% 
  ggscatter(x = "Order", y = "TIC", color = "Plate_Number") + 
  scale_color_manual(values = custom_colors, name = "Plate Number") +  # Map specific colors
  stat_cor() +ylim(0, 7e9)+ 
  ggtitle("TIC over Time for Samples")

# TIC over Time for Pooled Samples (a)
a <- data_tic_plate %>%
  dplyr::filter((str_detect(SampleID, regex("Pool", ignore_case = TRUE)))) %>%
  ggscatter(x = "Order", y = "TIC", color = "Plate_Number") + 
  scale_color_manual(values = custom_colors, name = "Plate Number") + 
  stat_cor() +ylim(0, 7e9)+ 
  ggtitle("TIC over Time for Pooled Samples")

# TIC over Time for 6mix Samples (b)
b <- data_tic_plate %>%
  dplyr::filter(str_detect(SampleID, "^QC_P[0-9]+_[0-9]+$") | str_detect(SampleID, regex("6mix", ignore_case = TRUE))) %>%
  ggscatter(x = "Order", y = "TIC", color = "Plate_Number") + 
  scale_color_manual(values = custom_colors, name = "Plate Number") + 
  stat_cor() +ylim(0, 7e9)+ 
  ggtitle("TIC over Time for 6mix Samples")

# TIC over Time for Blank Samples (c)
c <- data_tic_plate %>%
  dplyr::filter((str_detect(SampleID, regex("blank", ignore_case = TRUE)))) %>%
  ggscatter(x = "Order", y = "TIC", color = "Plate_Number") + 
  scale_color_manual(values = custom_colors, name = "Plate Number") + 
  stat_cor() +ylim(0, 7e9)+ 
  ggtitle("TIC over Time for Blank Samples")

# TIC over Time for Blank Swab Samples (CTL) (d)
d <- data_tic_plate %>%
  dplyr::filter((str_detect(SampleID, regex("CTL", ignore_case = TRUE)))) %>%
  ggscatter(x = "Order", y = "TIC", color = "Plate_Number") + 
  scale_color_manual(values = custom_colors, name = "Plate Number") + 
  stat_cor() +ylim(0, 7e9)+ 
  ggtitle("TIC over Time for Blank Swab Samples (CTL)")

f <- ggarrange(
  a0, a, b, c, d, 
  ncol = 3, nrow = 2, 
  labels = c("a", "b", "c", "d", "e"),
  common.legend = TRUE,  # Use a common legend
  legend = "bottom"      # Place the legend at the bottom
)

ggsave("../Figures/metabolomics_Figures/qc/TICvsOrder_all_platenumber.png", plot = last_plot(), dpi = 600, width = 16, height = 10, units = "in")
```

```{r}
# Filter by TIC 
# Convert Plate_Number to a factor for consistent coloring across all plots
data_tic_plate <- data_tic_plate %>%
  dplyr::mutate(Plate_Number = as.factor(Plate_Number))

# Define the mapping of specific colors to specific Plate_Number values
custom_colors <- c("1" = "#D55E00",  # Plate_Number == 1
                   "2" = "#2B579A",  # Plate_Number == 2
                   "3" = "#A1280B",  # Plate_Number == 3
                   "4" = "#217346")  # Plate_Number == 4

# TIC over Time for Samples (a0)
a0 <- data_tic_plate %>%
  dplyr::mutate(Plate_Number = as.factor(Plate_Number)) %>%
  dplyr::filter(!(str_detect(SampleID, regex("Blank|6mix|Pool|wash|CTL|QC", ignore_case = TRUE)))) %>%
  dplyr::filter(TIC >= 6e8 & TIC <= 5e9) %>%
  ggscatter(x = "Order", y = "TIC", color = "Plate_Number") + 
  scale_color_manual(values = custom_colors, name = "Plate Number") +  # Map specific colors
  stat_cor() +ylim(0, 7e9)+ 
  ggtitle("TIC over Time for Samples")

# TIC over Time for Pooled Samples (a)
a <- data_tic_plate %>%
  dplyr::filter((str_detect(SampleID, regex("Pool", ignore_case = TRUE)))) %>%
  ggscatter(x = "Order", y = "TIC", color = "Plate_Number") + 
  scale_color_manual(values = custom_colors, name = "Plate Number") + 
  stat_cor() +ylim(0, 7e9)+ 
  ggtitle("TIC over Time for Pooled Samples")

# TIC over Time for 6mix Samples (b)
b <- data_tic_plate %>%
  dplyr::filter(str_detect(SampleID, "^QC_P[0-9]+_[0-9]+$") | str_detect(SampleID, regex("6mix", ignore_case = TRUE))) %>%
  ggscatter(x = "Order", y = "TIC", color = "Plate_Number") + 
  scale_color_manual(values = custom_colors, name = "Plate Number") + 
  stat_cor() +ylim(0, 7e9)+ 
  ggtitle("TIC over Time for 6mix Samples")

# TIC over Time for Blank Samples (c)
c <- data_tic_plate %>%
  dplyr::filter((str_detect(SampleID, regex("blank", ignore_case = TRUE)))) %>%
  ggscatter(x = "Order", y = "TIC", color = "Plate_Number") + 
  scale_color_manual(values = custom_colors, name = "Plate Number") + 
  stat_cor() +ylim(0, 7e9)+ 
  ggtitle("TIC over Time for Blank Samples")

# TIC over Time for Blank Swab Samples (CTL) (d)
d <- data_tic_plate %>%
  dplyr::filter((str_detect(SampleID, regex("CTL", ignore_case = TRUE)))) %>%
  ggscatter(x = "Order", y = "TIC", color = "Plate_Number") + 
  scale_color_manual(values = custom_colors, name = "Plate Number") + 
  stat_cor() +ylim(0, 7e9)+ 
  ggtitle("TIC over Time for Blank Swab Samples (CTL)")

```
```{r}
f <- ggarrange(
  a0, a, b, c, d, 
  ncol = 3, nrow = 2, 
  labels = c("a", "b", "c", "d", "e"),
  common.legend = TRUE,  # Use a common legend
  legend = "bottom"      # Place the legend at the bottom
)

ggsave("../Figures/metabolomics_Figures/qc/TICvsOrder_all_platenumber_filtered.svg", plot = last_plot(), dpi = 600, width = 16, height = 10, units = "in")
```

```{r}
# Merging the `data_TIC` DataFrame with `metadata_2` to add metadata information based on matching `SampleID` in both DataFrames
data_tic_group <- data_TIC %>%
  left_join(metadata_2, by = c("SampleID" = "SampleID"))

# Define custom orange, pine green, brown, and Excel-style green color palette
custom_colors <- c("#df7963", "#67b2be", "#423fa6")  


# TIC over Time for Samples (a0)
a0 <- data_tic_group %>%
  dplyr::mutate(group = as.factor(category_label)) %>%
  dplyr::filter(!(str_detect(SampleID, regex("Blank|6mix|Pool|wash|CTL|QC", ignore_case = TRUE)))) %>%
  ggscatter(x = "Order", y = "TIC", color = "group") + 
  scale_color_manual(values = custom_colors, name = "group") +  # Use custom color palette
  stat_cor() + 
  ggtitle("TIC over Time for Samples")

ggsave("../Figures/metabolomics_Figures/qc/TICvsOrder_group.png", plot = last_plot(), dpi = 600, width = 8, height = 8, units = "in")
```

```{r}
# Merging the `data_TIC` DataFrame with `metadata_2` to add metadata information based on matching `SampleID` in both DataFrames
data_tic_group <- data_TIC %>%
  left_join(metadata_2, by = c("SampleID" = "SampleID"))

# Define custom orange, pine green, brown, and Excel-style green color palette
custom_colors <- c("#df7963", "#67b2be", "#423fa6")  

# TIC over Time for Samples (a0)
a0 <- data_tic_group %>%
  dplyr::mutate(group = as.factor(category_label)) %>%
  dplyr::filter(!(str_detect(SampleID, regex("Blank|6mix|Pool|wash|CTL|QC", ignore_case = TRUE)))) %>%
  ggscatter(x = "Order", y = "TIC", color = "group") + 
  scale_color_manual(values = custom_colors, name = "group") +  # Use custom color palette
  stat_cor() + 
  ggtitle("TIC over Time for Samples")

ggsave("../Figures/metabolomics_Figures/qc/TICvsOrder_group_filtered.png", plot = last_plot(), dpi = 600, width = 8, height = 8, units = "in")
```

```{r}
# Find the 6mix antibiotics
abundance_data <- subset(data_final_updated, SampleID == "QC_P2_6mix_4")

# Remove the SampleID column before transposing
abundance_data <- abundance_data[ , !(names(abundance_data) %in% "SampleID")]

# Transpose the data and convert values to numeric
abundance_data <- as.data.frame(t(abundance_data))
abundance_data <- as.numeric(as.matrix(abundance_data))

# Check internal standard and remove sample where there was a shift
fbmn_IS <- annotations %>% dplyr::filter(str_detect(Compound_Name, regex("sulf", ignore_case = TRUE))) %>% 
  distinct(X.Scan., .keep_all = TRUE) %>% dplyr::filter(Organism != "BILELIB19")

# Check internal standard and remove sample where there was a shift
fbmn_IS <- annotations %>% dplyr::filter(str_detect(Compound_Name, regex("Amitri", ignore_case = TRUE))) %>% 
  distinct(X.Scan., .keep_all = TRUE) %>% dplyr::filter(Organism != "BILELIB19")

# Extract IS feature from the table
table_IS <- data_final %>% column_to_rownames("SampleID") %>% t() %>% as.data.frame() %>% rownames_to_column("ID") %>%
  dplyr::filter(ID %in% fbmn_IS$X.Scan.) %>% column_to_rownames("ID") %>% t() %>% as.data.frame() %>%
  rownames_to_column("filename") %>% left_join(metadata_final, by = c("filename" = "SampleID")) %>%
  dplyr::filter(!(str_detect(filename, regex("Blank|6mix|Pool|wash|CTL|QC", ignore_case = TRUE))))# %>%
  #dplyr::select(filename, `15435`, Order, filename)
  # dplyr::select(filename, `15435`, Order, filename)

# Renaming the second column of `table_IS` to "sulfadimethoxine"
colnames(table_IS)[2] <- "sulfadimethoxine"

# Creating a scatter plot of "Order" vs. "sulfadimethoxine" with a regression line, correlation statistics, and setting y-axis limits
table_IS %>% 
  ggscatter(x = "Order", y = "sulfadimethoxine", add = c("reg.line")) +
  stat_cor() + # Adds correlation coefficient and p-value to the plot
  ylim(0, 5e6) # Sets y-axis limits from 0 to 5,000,000

ggsave("../Figures/metabolomics_Figures/qc/scatter_IS_filtered.png", plot = last_plot(), dpi = 600, width = 8, height = 6, units = "in")
```


```{r}
# Find the 6mix antibiotics
abundance_data <- subset(data_final_updated, SampleID == "QC_P2_6mix_4")

# Remove the SampleID column before transposing
abundance_data <- abundance_data[ , !(names(abundance_data) %in% "SampleID")]

# Transpose the data and convert values to numeric
abundance_data <- as.data.frame(t(abundance_data))
abundance_data <- as.numeric(as.matrix(abundance_data))

# Check internal standard and remove sample where there was a shift
fbmn_IS <- annotations %>% dplyr::filter(str_detect(Compound_Name, regex("sulf", ignore_case = TRUE))) %>% 
  distinct(X.Scan., .keep_all = TRUE) %>% dplyr::filter(Organism != "BILELIB19")

# Check internal standard and remove sample where there was a shift
fbmn_IS <- annotations %>% dplyr::filter(str_detect(Compound_Name, regex("Amitri", ignore_case = TRUE))) %>% 
  distinct(X.Scan., .keep_all = TRUE) %>% dplyr::filter(Organism != "BILELIB19")

# Extract IS feature from the table
table_IS <- data_final %>% column_to_rownames("SampleID") %>% t() %>% as.data.frame() %>% rownames_to_column("ID") %>%
  dplyr::filter(ID %in% fbmn_IS$X.Scan.) %>% column_to_rownames("ID") %>% t() %>% as.data.frame() %>%
  rownames_to_column("filename") %>% left_join(metadata_final, by = c("filename" = "SampleID")) %>%
  dplyr::filter(!(str_detect(filename, regex("Blank|6mix|Pool|wash|CTL|QC", ignore_case = TRUE))))# %>%
  #dplyr::select(filename, `15435`, Order, filename)
  # dplyr::select(filename, `15435`, Order, filename)

# Renaming the second column of `table_IS` to "sulfadimethoxine"
colnames(table_IS)[2] <- "sulfadimethoxine"

# Creating a scatter plot of "Order" vs. "sulfadimethoxine" with a regression line, correlation statistics, and setting y-axis limits
table_IS %>% 
  ggscatter(x = "Order", y = "sulfadimethoxine", add = c("reg.line")) +
  stat_cor() + # Adds correlation coefficient and p-value to the plot
  ylim(0, 5e6) # Sets y-axis limits from 0 to 5,000,000

ggsave("../Figures/metabolomics_Figures/qc/scatter_IS_filtered.png", plot = last_plot(), dpi = 600, width = 8, height = 6, units = "in")
```


Scan Numbers of the 6 antibiotics (Sulfadimethoxine was the one that was used according to Jasmine )
10110 Sulfamethazine
15435 Sulfadimethoxine
16885 Amitriptyline
21848 Coumarin
12511 Sulfachloropyridazine
10762 Sulfadimethizole

```{r}
# Make plot for all 6-mix antibiotics

colnames(table_IS)[2] <- "Coumarin_314"

Coumarin <- table_IS %>% ggscatter(x = "Order", y = "Coumarin_314", add = c("reg.line")) +
  stat_cor() + ylim(0, 7e9)#+ ylim(0, 4e7)

cv_is <- sd(table_IS$Coumarin_314)/mean(table_IS$Coumarin_314)
#24.75%

colnames(table_IS)[2] <- "Sulfachloropyridazine"

Sulfachloropyridazine <- table_IS %>% ggscatter(x = "Order", y = "Sulfachloropyridazine", add = c("reg.line")) +
  stat_cor() + ylim(0, 7e9)#+ ylim(0, 4e7)

cv_is <- sd(table_IS$Sulfachloropyridazine)/mean(table_IS$Sulfachloropyridazine)
#66.78%

colnames(table_IS)[2] <- "Sulfadimethizole"

Sulfadimethizole <- table_IS %>% ggscatter(x = "Order", y = "Sulfadimethizole", add = c("reg.line")) +
  stat_cor() + ylim(0, 7e9)#+ ylim(0, 4e7)

cv_is <- sd(table_IS$Sulfadimethizole)/mean(table_IS$Sulfadimethizole)
#68.34%

colnames(table_IS)[2] <- "Amitriptyline"

Amitriptyline <- table_IS %>% ggscatter(x = "Order", y = "Amitriptyline", add = c("reg.line")) +
  stat_cor()  + ylim(0, 7e9)#+ ylim(0, 4e7)

cv_is <- sd(table_IS$Amitriptyline)/mean(table_IS$Amitriptyline)
#29.3%

colnames(table_IS)[2] <- "Sulfadimethoxine"

sulfadimethoxine <- table_IS %>% ggscatter(x = "Order", y = "Sulfadimethoxine", add = c("reg.line")) +
  stat_cor() + ylim(0, 7e9)#+ ylim(0, 4e7)

cv_is <- sd(table_IS$Sulfadimethoxine)/mean(table_IS$Sulfadimethoxine)
#60.5%

#10110
colnames(table_IS)[2] <- "Sulfamethazine"

sulfamethazine <- table_IS %>% ggscatter(x = "Order", y = "Sulfamethazine", add = c("reg.line")) +
  stat_cor() + ylim(0, 7e9)#ylim(0, 4e7)
cv_is <- sd(table_IS$Sulfamethazine)/mean(table_IS$Sulfamethazine)
#72.5%

f <- ggarrange(
  sulfamethazine, sulfadimethoxine, Amitriptyline, Sulfadimethizole, Sulfachloropyridazine, Coumarin,
  ncol = 3, nrow = 2, 
  labels = c("a", "b", "c", "d", "e", "f"),
  common.legend = TRUE,  # Use a common legend
  legend = "bottom"      # Place the legend at the bottom
)

ggsave("../Figures/metabolomics_Figures/qc/sixmix_antibiotics_7e9.png", plot = last_plot(), dpi = 600, width = 12, height = 8, units = "in")

ggsave("../Figures/metabolomics_Figures/qc/PeakArea_IS_sulfamethazine.png", plot = last_plot(), dpi = 600, width = 8, height = 6, units = "in")

ggsave("../Figures/metabolomics_Figures/qc/PeakArea_IS_filtered.png", plot = last_plot(), dpi = 600, width = 8, height = 6, units = "in")
```

```{r}
# Filter table_IS for filenames with sulfadimethoxin < 22378
filtered_table_IS <- table_IS[table_IS$sulfadimethoxin < 1e6, "filename"]

# Filter data_TIC for SampleIDs with TIC < 569590420
filtered_data_TIC <- data_TIC[data_TIC$TIC < 5e8, "SampleID"]

# Find common values
common_filenames <- intersect(filtered_table_IS, filtered_data_TIC)

# Find values in table_IS but not in data_TIC
unique_to_table_IS <- setdiff(filtered_table_IS, filtered_data_TIC)

# Find values in data_TIC but not in table_IS
unique_to_data_TIC <- setdiff(filtered_data_TIC, filtered_table_IS)

# Print results
cat("Common filenames:\n")
print(common_filenames)

cat("\nFilenames in table_IS but not in data_TIC:\n")
print(unique_to_table_IS)

cat("\nFilenames in data_TIC but not in table_IS:\n")
print(unique_to_data_TIC)
```


```{r}
# Check features in blanks and pools
data_blank <- data_final %>% dplyr::filter((str_detect(SampleID, regex("blank", ignore_case = TRUE)))) 
data_pools <- data_final %>% dplyr::filter((str_detect(SampleID, regex("Pool", ignore_case = TRUE))))
data_sixmix <- data_final %>% dplyr::filter(str_detect(SampleID, "^QC_P[0-9]+_[0-9]+$") | str_detect(SampleID, regex("6mix", ignore_case = TRUE)))
data_CTL <- data_final %>% dplyr::filter((str_detect(SampleID, regex("CTL", ignore_case = TRUE))))

# Check features in blanks and pools with filtering out the outliers
data_blank <- data_final_updated %>% dplyr::filter((str_detect(SampleID, regex("blank", ignore_case = TRUE)))) 
data_pools <- data_final_updated %>% dplyr::filter((str_detect(SampleID, regex("Pool", ignore_case = TRUE))))
data_sixmix <- data_final_updated %>% dplyr::filter(str_detect(SampleID, "^QC_P[0-9]+_[0-9]+$") | str_detect(SampleID, regex("6mix", ignore_case = TRUE)))
data_CTL <- data_final_updated %>% dplyr::filter((str_detect(SampleID, regex("CTL", ignore_case = TRUE))))

# Get feature info for blanks
blanks_feature_info <- data.frame(Feature = colnames(data_blank)[-1],
                                  Mean_blank = data_blank %>% column_to_rownames("SampleID") %>% colMeans(), 
                                  SD_blank =  data_blank %>% column_to_rownames("SampleID") %>% apply(2, sd)) %>%
  dplyr::mutate(CV_blank = SD_blank/Mean_blank) %>% left_join(annotations, by = c("Feature" = "X.Scan.")) %>% 
  left_join(info_feature) %>% dplyr::select(Feature, mz, RT, Compound_Name, SpectrumID, 
                                            Precursor_MZ, Mean_blank, SD_blank, CV_blank) %>% 
  dplyr::filter(Mean_blank > 0) %>% arrange(desc(Mean_blank))

# Get feature info for six mix
sixmix_feature_info <- data.frame(Feature = colnames(data_sixmix)[-1],
                                  Mean_sixmix = data_sixmix %>% column_to_rownames("SampleID") %>% colMeans(), 
                                  SD_sixmix = data_sixmix %>% column_to_rownames("SampleID") %>% apply(2, sd)) %>%
  dplyr::mutate(CV_sixmix = SD_sixmix/Mean_sixmix) %>% left_join(annotations, by = c("Feature" = "X.Scan.")) %>% 
  left_join(info_feature) %>% dplyr::select(Feature, mz, RT, Compound_Name, SpectrumID, 
                                            Precursor_MZ, Mean_sixmix, SD_sixmix, CV_sixmix) %>% 
  dplyr::filter(Mean_sixmix > 0) %>% arrange(desc(Mean_sixmix))

# Get feature info for pool
pools_feature_info <- data.frame(Feature = colnames(data_pools)[-1],
                                 Mean_pool = data_pools %>% column_to_rownames("SampleID") %>% colMeans(), 
                                 SD_pool =  data_pools %>% column_to_rownames("SampleID") %>% apply(2, sd)) %>%
  dplyr::mutate(CV_pool = SD_pool/Mean_pool) %>% left_join(annotations, by = c("Feature" = "X.Scan.")) %>% 
  left_join(info_feature) %>% dplyr::select(Feature, mz, RT, Compound_Name, SpectrumID, 
                                            Precursor_MZ, Mean_pool, SD_pool, CV_pool) %>% 
  dplyr::filter(Mean_pool > 0) %>% arrange(desc(Mean_pool))

# Get feature info for CTLs
CTLs_feature_info <- data.frame(Feature = colnames(data_CTL)[-1],
                                 Mean_ctl = data_CTL %>% column_to_rownames("SampleID") %>% colMeans(), 
                                 SD_ctl =  data_CTL %>% column_to_rownames("SampleID") %>% apply(2, sd)) %>%
  dplyr::mutate(CV_ctl = SD_ctl/Mean_ctl) %>% left_join(annotations, by = c("Feature" = "X.Scan.")) %>% 
  left_join(info_feature) %>% dplyr::select(Feature, mz, RT, Compound_Name, SpectrumID, 
                                            Precursor_MZ, Mean_ctl, SD_ctl, CV_ctl) %>% 
  dplyr::filter(Mean_ctl > 0) %>% arrange(desc(Mean_ctl))
```

```{r}
# Features to be removed Pools/Blank < 5 or 10
feature_to_remove <- blanks_feature_info %>% left_join(pools_feature_info) %>%
  dplyr::filter(Mean_blank > 0) %>% 
  dplyr::mutate(Pool_Blank = Mean_pool/Mean_blank) %>% 
  dplyr::filter(Pool_Blank < 5 | is.na(Pool_Blank))

# Data with blank removal
data_clean <- data_final %>% dplyr::select(-c(feature_to_remove$Feature)) %>%
  column_to_rownames("SampleID") %>%
  select_if(~sum(.) != 0) %>% rownames_to_column("SampleID")

# Data with blank removal after outlier removal
data_clean <- data_final_updated %>% dplyr::select(-c(feature_to_remove$Feature)) %>%
  column_to_rownames("SampleID") %>%
  select_if(~sum(.) != 0) %>% rownames_to_column("SampleID")

# Features to be removed Pool/QCmix < 5 or 10 
feature_to_remove_mix <- sixmix_feature_info %>% left_join(pools_feature_info) %>% 
  dplyr::filter(Mean_sixmix > 0) %>% 
  dplyr::mutate(Pool_Mix = Mean_pool/Mean_sixmix) %>% 
  dplyr::filter(Pool_Mix < 5 | is.na(Pool_Mix)) %>% dplyr::filter(!(Feature %in% feature_to_remove$Feature))

# Data with blank2 removal
data_clean2 <- data_clean %>% dplyr::select(-c(feature_to_remove_mix$Feature))

# Features to be removed Pool/CTL < 5 or 10 
feature_to_remove_ctl <- CTLs_feature_info %>% left_join(pools_feature_info) %>% 
  dplyr::filter(Mean_ctl > 0) %>% 
  dplyr::mutate(Pool_CTL = Mean_pool/Mean_ctl) %>% 
  dplyr::filter(Pool_CTL < 5 | is.na(Pool_CTL)) %>% dplyr::filter(!(Feature %in% feature_to_remove_mix$Feature))  %>%  dplyr::filter(!(Feature %in% feature_to_remove$Feature))

# Data with blank2 removal
data_clean3 <- data_clean2 %>% dplyr::select(-c(feature_to_remove_ctl$Feature))
```