---
title: "Metabolomics_Loreal"
output: html_document
date: "2024-09-25"
---

```{r packages}
library(tidyverse)
library(mixOmics)
library(ggpubr)
library(vegan)
library(caret)
library(rstatix)
library(UpSetR)
library(homologueDiscoverer)
library(ggdensity)
library(cowplot)
library(readxl)
library(lmerTest)
library(RColorBrewer)
library(MassTools)
library(shiny)
library(ggplot2)
library(dplyr)
library(stringr)
library(svglite)
library(Cairo)
```
```{r}
#Function for homologues analysis
plotAnnotatedStatic <- function(annotated, legend_setting = "bottom"){
  
  annotated <- mutate(annotated,
                      homologue_id = if_else(is.na(homologue_id), 0L, homologue_id),
                      homologue_id = as.factor(homologue_id))
  colourCount = length(unique(annotated$homologue_id))
  getPalette = colorRampPalette(RColorBrewer::brewer.pal(9, "Set1"))
  ncolor = length(unique(annotated$homologue_id))
  annotated <- arrange(annotated, desc(homologue_id))
  g <- ggplot(annotated, aes(group = homologue_id)) +
    geom_point(aes(x = rt, y = mz, color = homologue_id, size = homologue_id,
                   shape = homologue_id, alpha = homologue_id)) +
    geom_line(data = filter(annotated, homologue_id != 0),
              aes(x = rt, y = mz, group = homologue_id), alpha = 0.4, linewidth = 0.1) +
    ggtitle("Annotated Peak Table") +
    scale_colour_manual(values = c("black", getPalette(colourCount-1))) +
    scale_alpha_manual(values =  c(0.2, rep(0.85, times = ncolor))) +
    scale_size_manual(values = c(0.1, rep(1.5, times = ncolor))) +
    scale_shape_manual(values = c(1, rep(19, times = ncolor))) +
    xlab("Retention Time (s)") +
    ylab("Mass to Charge Ratio") +
    theme(legend.position=legend_setting, text = element_text(family="mono"))
  return(g)
  
}
```



```{r}
#Run 3
feature_table <- read_csv("../Data/metabolomics/Run3_10252024/FBMN_mgf_Lami_10252024_quant.csv")

metadata <- read_csv("../Metadata/metadata_final_18062024.csv")

sample_order <- read.csv("../Data/metabolomics/Run3_10252024/Sequence_order_10262024.csv")

annotations <- read.delim("../Data/metabolomics/Run3_10252024/merged_results_with_gnps.tsv")
annotations$X.Scan. <- as.character(annotations$X.Scan.)
#canopus <- read_tsv("../data/canopus_compound_summary.tsv")
#canopus$featureId <- as.character(canopus$featureId)

mapping_file <- read.csv("../Data/metabolomics/Run3_10252024/Mapping_10262024.csv")
```


```{r}
mapping_file$Manifest_ID_Sample_Name <- gsub("_", ".", mapping_file$Manifest_ID_Sample_Name)

metadata_2 <- metadata %>% left_join(mapping_file, by = c("sample_name" = "Manifest_ID_Sample_Name"))
```

```{r}
metadata_2$SampleID <- metadata_2$Filename_ID
```
```{r}
# Reorder the columns, placing metadata_final$SampleID as the first column
metadata_2 <- metadata_2[, c("SampleID", setdiff(names(metadata_2), "SampleID"))]

```


```{r}
# Find the rows in mapping_file where Filename_ID is not present in metadata_final$SampleID
missing_sample_ids <- mapping_file[!(mapping_file$Filename_ID %in% metadata_2$SampleID), ]

# Create a new data frame with only the missing Filename_ID as SampleID
missing_sample_ids_df <- data.frame(SampleID = missing_sample_ids$Filename_ID)

# Append only the SampleIDs to metadata_final
# This assumes metadata_final has more columns, and we just add the missing SampleIDs with other columns left blank
missing_rows <- data.frame(SampleID = missing_sample_ids_df$SampleID, 
                           matrix(NA, nrow = nrow(missing_sample_ids_df), ncol = ncol(metadata_2) - 1))
colnames(missing_rows) <- colnames(metadata_2)

# Now append the missing rows with SampleID and NA for other columns
metadata_final <- rbind(metadata_2, missing_rows)

# Print or save the updated metadata_final
head(metadata_final)


```
```{r}
# Assuming updated_metadata_final is your data frame with SampleID column

metadata_final <- metadata_final %>%
  mutate(Sample_Type = case_when(
    str_detect(SampleID, regex("wash", ignore_case = TRUE)) ~ "Wash",
    str_detect(SampleID, regex("Blank", ignore_case = TRUE)) ~ "Blank",
    str_detect(SampleID, "^QC_P[0-9]+_[0-9]+$") | str_detect(SampleID, regex("6mix", ignore_case = TRUE)) ~ "6mix",
    str_detect(SampleID, regex("Pool", ignore_case = TRUE)) ~ "Pooled samples",
    str_detect(SampleID, regex("CTL", ignore_case = TRUE)) ~ "CTL",
    TRUE ~ "samples"  # Default to 'samples' for SampleIDs that don't match the others
  ))

# Print or inspect the updated metadata_final with Sample_Type column
head(metadata_final)
```



```{r}
sum(metadata_final$Sample_Type=="samples")
```
```{r}
sum(metadata_final$Sample_Type=="6mix")
```
```{r}
sum(metadata$group=="Healthy")
```

```{r}
sum(metadata$group=="Acne_L")
```
```{r}
sum(metadata$group=="Acne_NL")
```

```{r}
#L'Oreal data
metadata_final <- metadata_final %>% left_join(sample_order, by = c("SampleID" = "File.Name"))

```

```{r}
unmatched_metadata <- metadata_final %>%
  anti_join(sample_order, by = c("SampleID" = "File.Name"))

# View the result
unmatched_metadata

```
```{r}
unmatched_sample_order <- sample_order %>%
  anti_join(metadata_final, by = c("File.Name" = "SampleID"))

# View the result
unmatched_sample_order

```

```{r}
# Check for duplicated SampleIDs after the join
duplicated_samples <- metadata_final %>%
  group_by(SampleID) %>%
  filter(n() > 1)

# Display the duplicated SampleIDs
cat("Duplicated SampleIDs after the join:\n")
print(duplicated_samples)

# Optionally, count the number of duplicates
num_duplicates <- nrow(duplicated_samples)
cat("\nNumber of duplicated SampleIDs after the join: ", num_duplicates, "\n")
```

```{r}

info_feature <- feature_table %>% dplyr::select(1:3,7)
colnames(info_feature) <- c("Feature", "mz", "RT", "Corr_ID")
info_feature$Feature <- as.character(info_feature$Feature)

info_feature_complete <- info_feature %>% 
  left_join(annotations, by = c("Feature" = "X.Scan.")) %>%
  dplyr::select(1:4,18,24) %>% 
  left_join(canopus %>% distinct(featureId, .keep_all = TRUE), by = c("Feature" = "featureId"))



```




```{r}
# Extract data table 
data <- feature_table %>% 
  column_to_rownames("row ID") %>% dplyr::select(contains("Peak")) %>% 
  t() %>% as.data.frame() %>% rownames_to_column("SampleID") %>% 
  arrange(SampleID) %>% distinct(SampleID, .keep_all = TRUE)

data$SampleID <- gsub(".mzML Peak area", "", data$SampleID)


```
```{r}
# Find the rows in mapping_file where Filename_ID is not present in metadata_final$SampleID
missing_sample_ids <- data[!(data$SampleID %in% metadata_final$SampleID), ]

```

```{r}
# Find the rows in mapping_file where Filename_ID is not present in metadata_final$SampleID
missing_samples <- metadata_final[!(metadata_final$SampleID %in% data$SampleID), ]
```
```{r}
# Assuming your metadata_final and data are data frames with a column named "SampleID"
# Replace 'metadata_final' and 'data' with your actual data frame names
metadata_ids <- metadata_final$SampleID
data_ids <- data$SampleID

# 1. Find SampleIDs in metadata_final but not in data
missing_in_data <- setdiff(metadata_ids, data_ids)

# 2. Find SampleIDs in data but not in metadata_final
missing_in_metadata <- setdiff(data_ids, metadata_ids)

# 3. Check for duplicated SampleIDs in metadata_final
duplicates_in_metadata <- metadata_ids[duplicated(metadata_ids)]

# 4. Check for duplicated SampleIDs in data
duplicates_in_data <- data_ids[duplicated(data_ids)]

# Print results
cat("SampleIDs in metadata_final but not in data:\n")
print(missing_in_data)

cat("\nSampleIDs in data but not in metadata_final:\n")
print(missing_in_metadata)

cat("\nDuplicated SampleIDs in metadata_final:\n")
print(duplicates_in_metadata)

cat("\nDuplicated SampleIDs in data:\n")
print(duplicates_in_data)

```
```{r}
# Discard first blanks, pools and mix for conditioning L_Oreal Table
data_final <- data %>% 
  dplyr::filter(!(SampleID %in% c("blank_P2_0A_prerun", "blank_P3_0A_prerun", "blank_P3_0B_prerun","blank_P1_0A_prerun","QC_P1_6mix_0", "QC_P1_Pool_0")))



```
```{r}
# List of samples to discard
samples_to_remove <- c("blank_P2_0A_prerun", "blank_P3_0A_prerun", "blank_P3_0B_prerun",
                       "blank_P1_0A_prerun", "QC_P1_6mix_0", "QC_P1_Pool_0")

# Check if any sample is missing in the dataset
missing_samples <- setdiff(samples_to_remove, data$SampleID)

# Issue a warning for missing samples
if (length(missing_samples) > 0) {
  warning(paste("The following samples are not present in the dataset:", paste(missing_samples, collapse = ", ")))
}

```
```{r}
data_final <- data %>% 
  dplyr::filter(!(SampleID %in% c("blank_P2_0A_prerun", "blank_P3_0A_prerun_20241022011201", "blank_P3_0B_prerun","blank_P1_0A_prerun","QC_P1_6mix_0", "QC_P1_Pool_0")))
```
```{r}
samples_to_remove <- c("blank_P2_0A_prerun", "blank_P3_0A_prerun_20241022011201", "blank_P3_0B_prerun",
                       "blank_P1_0A_prerun", "QC_P1_6mix_0", "QC_P1_Pool_0")

# Check if any sample is missing in the dataset
missing_samples <- setdiff(samples_to_remove, data$SampleID)

# Issue a warning for missing samples
if (length(missing_samples) > 0) {
  warning(paste("The following samples are not present in the dataset:", paste(missing_samples, collapse = ", ")))
}

```

```{r}
# Calculate Total Ion Count (TIC)
data_TIC <- data.frame(TIC = rowSums(data_final %>% column_to_rownames("SampleID"))) %>% 
  rownames_to_column("SampleID") %>% left_join(sample_order, by = c("SampleID" = "File.Name"))


```
```{r}
#Plots a scatter plot of the Order vs. TIC
data_TIC %>% dplyr::filter(!(str_detect(SampleID, regex("Blank|6mix|Pool|wash|CTL|QC", ignore_case = TRUE)))) %>%
  left_join(metadata_final) %>%
  ggscatter("Order", "TIC", color = "subject_ID", add = "reg.line") +
  stat_cor()
```
```{r}
ggsave("../figures/sequence_order/TICvsOrder.png", plot = last_plot(), dpi = 300, width = 8, height = 6, units = "in")

```
data_combined <- data_final %>%
  dplyr::filter(str_detect(SampleID, "^QC_P[0-9]+_[0-9]+$") | 
                str_detect(SampleID, regex("6mix", ignore_case = TRUE)))
```{r}
ggsave("../figures/sequence_order/TICvsOrder.svg", plot = last_plot(), dpi = 300, width = 8, height = 6, units = "in", device = "svg")


```


```{r}
# Check sample type
sample_tic <-  data_TIC %>% dplyr::filter(!(str_detect(SampleID, regex("Blank|6mix|Pool|wash|CTL|QC", ignore_case = TRUE)))) %>% summarise(mean(TIC))
pool_tic <- data_TIC %>% dplyr::filter((str_detect(SampleID, regex("Pool", ignore_case = TRUE)))) %>% summarise(mean(TIC))
six_tic <- data_TIC %>%   dplyr::filter(str_detect(SampleID, "^QC_P[0-9]+_[0-9]+$") | str_detect(SampleID, regex("6mix", ignore_case = TRUE))) %>% summarise(mean(TIC))
blank_tic <- data_TIC %>% dplyr::filter((str_detect(SampleID, regex("blank", ignore_case = TRUE)))) %>% summarise(mean(TIC))
#internal control swabs given by L'Oréal
ctl_tic <- data_TIC %>% dplyr::filter((str_detect(SampleID, regex("CTL", ignore_case = TRUE)))) %>% summarise(mean(TIC))


ratio_tic_pb <- pool_tic/sample_tic
ratio_tic_sb <- sample_tic/blank_tic
ratio_tic_6p <- pool_tic/six_tic
```



```{r}
# Check TIC overtime for QCpool and Blanks
a0 <- data_TIC %>% dplyr::filter(!(str_detect(SampleID, regex("Blank|6mix|Pool|wash|CTL|QC", ignore_case = TRUE)))) %>% 
  ggscatter("Order", "TIC", add = "reg.line") +
  stat_cor() + ggtitle("TIC over Time for Samples")

a <- data_TIC %>% dplyr::filter((str_detect(SampleID, regex("Pool", ignore_case = TRUE))))%>% 
  ggscatter("Order", "TIC", add = "reg.line") +
  stat_cor() + ggtitle("TIC over Time for Pooled Samples")

b <- data_TIC %>%   dplyr::filter(str_detect(SampleID, "^QC_P[0-9]+_[0-9]+$") | str_detect(SampleID, regex("6mix", ignore_case = TRUE)))  %>% 
  ggscatter("Order", "TIC", add = "reg.line") +
  stat_cor() + ggtitle("TIC over Time for 6mix Samples")

c <- data_TIC %>% dplyr::filter((str_detect(SampleID, regex("blank", ignore_case = TRUE)))) %>% 
  ggscatter("Order", "TIC", add = "reg.line") +
  stat_cor()  + ggtitle("TIC over Time for Blank Samples")

d <- data_TIC %>%  dplyr::filter((str_detect(SampleID, regex("CTL", ignore_case = TRUE)))) %>% 
  ggscatter("Order", "TIC", add = "reg.line") +
  stat_cor() + ggtitle("TIC over Time for Blank Swab Samples (CTL)")



```

```{r}
f <- ggarrange(a0, a, b, c, d, ncol = 3, nrow = 2, labels = c("a", "b", "c", "d", "e"))

```
##RUN1: no blanks since blanks were already subtracted!
```{r}
# Check TIC overtime for QCpool and Blanks RUN1
a0 <- data_TIC %>% dplyr::filter(!(str_detect(SampleID, regex("Blank|6mix|Pool|wash|CTL|QC", ignore_case = TRUE)))) %>% 
  ggscatter("Order", "TIC", add = "reg.line") +
  stat_cor() + ggtitle("TIC over Time for Samples")

a <- data_TIC %>% dplyr::filter((str_detect(SampleID, regex("COSM", ignore_case = TRUE))))%>% 
  ggscatter("Order", "TIC", add = "reg.line") +
  stat_cor() + ggtitle("TIC over Time for COSM Control Samples")

b <- data_TIC %>%   dplyr::filter(str_detect(SampleID, "^QC_P[0-9]+_[0-9]+$") | str_detect(SampleID, regex("6mix", ignore_case = TRUE)))  %>% 
  ggscatter("Order", "TIC", add = "reg.line") +
  stat_cor() + ggtitle("TIC over Time for 6mix Samples")


d <- data_TIC %>%  dplyr::filter((str_detect(SampleID, regex("CTL", ignore_case = TRUE)))) %>% 
  ggscatter("Order", "TIC", add = "reg.line") +
  stat_cor() + ggtitle("TIC over Time for Blank Swab Samples (CTL)")


```
```{r}
f <- ggarrange(a0, a, b, d, ncol = 2, nrow = 2, labels = c("a", "b", "c", "d"))

```
```{r}
ggsave("../figures/TICvsOrder_all_RUN1.png", plot = last_plot(), dpi = 300, width = 12, height = 10, units = "in")
```

```{r}
ggsave("../figures/sequence_order/TICvsOrder_all.png", plot = last_plot(), dpi = 300, width = 16, height = 10, units = "in")

```

```{r}
#to plot the Plate_Number run this code 
#sample_order_2 <- read.csv("../data/rerun_blue_sequence_p1_4.csv") %>% dplyr::select(1,3,23)
```

```{r}
#to plot the Plate_Number run this code 
sample_order_2 <- read.csv("../data/Sequence_Rerun_09272024.csv") %>% dplyr::select(1,3,23)
```

```{r}
data_tic_plate <- data_TIC %>%
  left_join(sample_order_2, by = c("SampleID" = "File.Name", "Order" = "Order"))

```
```{r}
# SampleIDs in data_TIC that did not merge into data_tic_plate (i.e., not present in sample_order_2)
not_in_sample_order <- data_TIC %>%
  anti_join(sample_order_2, by = c("SampleID" = "File.Name", "Order" = "Order"))

# SampleIDs in sample_order_2 that did not merge into data_tic_plate (i.e., not present in data_TIC)
not_in_data_TIC <- sample_order_2 %>%
  anti_join(data_TIC, by = c("File.Name" = "SampleID", "Order" = "Order"))

```



```{r}
# Convert Plate_Number to a factor for consistent coloring across all plots
data_tic_plate <- data_tic_plate %>%
  dplyr::mutate(Plate_Number = as.factor(Plate_Number))

# Define the mapping of specific colors to specific Plate_Number values
custom_colors <- c("1" = "#D55E00",  # Plate_Number == 1
                   "2" = "#2B579A",  # Plate_Number == 2
                   "3" = "#A1280B",  # Plate_Number == 3
                   "4" = "#217346")  # Plate_Number == 4

# TIC over Time for Samples (a0)
a0 <- data_tic_plate %>%
  dplyr::mutate(Plate_Number = as.factor(Plate_Number)) %>%
  dplyr::filter(!(str_detect(SampleID, regex("Blank|6mix|Pool|wash|CTL|QC", ignore_case = TRUE)))) %>%
  ggscatter(x = "Order", y = "TIC", color = "Plate_Number") + 
  scale_color_manual(values = custom_colors, name = "Plate Number") +  # Map specific colors
  stat_cor() + 
  ggtitle("TIC over Time for Samples")

# TIC over Time for Pooled Samples (a)
a <- data_tic_plate %>%
  dplyr::filter((str_detect(SampleID, regex("Pool", ignore_case = TRUE)))) %>%
  ggscatter(x = "Order", y = "TIC", color = "Plate_Number") + 
  scale_color_manual(values = custom_colors, name = "Plate Number") + 
  stat_cor() + 
  ggtitle("TIC over Time for Pooled Samples")

# TIC over Time for 6mix Samples (b)
b <- data_tic_plate %>%
  dplyr::filter(str_detect(SampleID, "^QC_P[0-9]+_[0-9]+$") | str_detect(SampleID, regex("6mix", ignore_case = TRUE))) %>%
  ggscatter(x = "Order", y = "TIC", color = "Plate_Number") + 
  scale_color_manual(values = custom_colors, name = "Plate Number") + 
  stat_cor() + 
  ggtitle("TIC over Time for 6mix Samples")

# TIC over Time for Blank Samples (c)
c <- data_tic_plate %>%
  dplyr::filter((str_detect(SampleID, regex("blank", ignore_case = TRUE)))) %>%
  ggscatter(x = "Order", y = "TIC", color = "Plate_Number") + 
  scale_color_manual(values = custom_colors, name = "Plate Number") + 
  stat_cor() + 
  ggtitle("TIC over Time for Blank Samples")

# TIC over Time for Blank Swab Samples (CTL) (d)
d <- data_tic_plate %>%
  dplyr::filter((str_detect(SampleID, regex("CTL", ignore_case = TRUE)))) %>%
  ggscatter(x = "Order", y = "TIC", color = "Plate_Number") + 
  scale_color_manual(values = custom_colors, name = "Plate Number") + 
  stat_cor() + 
  ggtitle("TIC over Time for Blank Swab Samples (CTL)")


```

```{r}
f <- ggarrange(
  a0, a, b, c, d, 
  ncol = 3, nrow = 2, 
  labels = c("a", "b", "c", "d", "e"),
  common.legend = TRUE,  # Use a common legend
  legend = "bottom"      # Place the legend at the bottom
)

```
```{r}
ggsave("../figures/sequence_order/TICvsOrder_all_platenumber.png", plot = last_plot(), dpi = 300, width = 16, height = 10, units = "in")

```

```{r}
ggsave("../figures/sequence_order/TICvsOrder_all_platenumber.svg", plot = last_plot(), dpi = 300, width = 16, height = 10, units = "in")
```

```{r}

data_tic_group <- data_TIC %>%
  left_join(metadata_2, by = c("SampleID" = "SampleID"))

# Define custom orange, pine green, brown, and Excel-style green color palette
custom_colors <- c("#D55E00", "#2B579A", "#A1280B")  


# TIC over Time for Samples (a0)
a0 <- data_tic_group %>%
  dplyr::mutate(group = as.factor(group)) %>%
  dplyr::filter(!(str_detect(SampleID, regex("Blank|6mix|Pool|wash|CTL|QC", ignore_case = TRUE)))) %>%
  ggscatter(x = "Order", y = "TIC", color = "group") + 
  scale_color_manual(values = custom_colors, name = "group") +  # Use custom color palette
  stat_cor() + 
  ggtitle("TIC over Time for Samples")


```

```{r}
ggsave("../figures/sequence_order/TICvsOrder_group.png", plot = last_plot(), dpi = 300, width = 8, height = 8, units = "in")
```

```{r}
data_tic_group3 <- data_TIC %>%
  left_join(metadata_filtered, by = c("SampleID" = "SampleID")) %>%
  filter(!Plate_Number == 2) 

# Define custom orange, pine green, brown, and Excel-style green color palette
custom_colors <- c("#D55E00", "#2B579A", "#A1280B")  


# TIC over Time for Samples (a0)
a0 <- data_tic_group3 %>%
  dplyr::mutate(group = as.factor(group)) %>%
  dplyr::filter(!(str_detect(SampleID, regex("Blank|6mix|Pool|wash|CTL|QC", ignore_case = TRUE)))) %>%
  ggscatter(x = "Order.x", y = "TIC", color = "group") + 
  scale_color_manual(values = custom_colors, name = "group") +  # Use custom color palette
  stat_cor() + 
  ggtitle("TIC over Time for Samples")

```

```{r}
ggsave("../figures/sequence_order/TICvsOrder_group_filtered.png", plot = last_plot(), dpi = 300, width = 8, height = 8, units = "in")
```

```{r}
# Check internal standard and remove sample where there was a shift
fbmn_IS <- annotations %>% dplyr::filter(str_detect(Compound_Name, regex("sulf", ignore_case = TRUE))) %>% 
  distinct(X.Scan., .keep_all = TRUE) %>% dplyr::filter(Organism != "BILELIB19")
```


```{r}
#sulfadimethoxine was the one that was used according to Jasmine 
#instead of 974 --> X.Scan 18243
# Extract IS feature from the table
table_IS <- data_final %>% column_to_rownames("SampleID") %>% t() %>% as.data.frame() %>% rownames_to_column("ID") %>% 
  dplyr::filter(ID %in% fbmn_IS$X.Scan.) %>% column_to_rownames("ID") %>% t() %>% as.data.frame() %>% 
  rownames_to_column("filename") %>% left_join(metadata_final, by = c("filename" = "SampleID")) %>%
  dplyr::filter(!(str_detect(filename, regex("Blank|6mix|Pool|wash|CTL|QC", ignore_case = TRUE)))) %>%
  dplyr::select(filename, `18243`, Order, filename)
```


```{r}
colnames(table_IS)[2] <- "sulfadimethoxine"

table_IS %>% ggscatter(x = "Order", y = "sulfadimethoxine", add = c("reg.line")) +
  stat_cor()
```

```{r}
ggsave("../figures/sequence_order/scatter_IS.png", plot = last_plot(), dpi = 300, width = 8, height = 6, units = "in")
```

```{r}
ggsave("../figures/sequence_order/scatter_IS.svg", plot = last_plot(), dpi = 300, width = 8, height = 6, units = "in")
```

```{r}
table_IS %>% ggbarplot(x = "Order", y = "sulfadimethoxine", xlab = "Run Order", 
                       ylab = "Peak Area sulfadimethoxine", title = "Internal Standard Acquisition") +
  geom_hline(yintercept = mean(table_IS$sulfadimethoxine, na.rm = TRUE), linetype = "dashed", color = "blue")

cv_is <- sd(table_IS$sulfadimethoxine)/mean(table_IS$sulfadimethoxine)

#CV is 37,57% 
```

```{r}
ggsave("../figures/sequence_order/PeakArea_IS.png", plot = last_plot(), dpi = 300, width = 8, height = 6, units = "in")
```
```{r}
ggsave("../figures/sequence_order/PeakArea_IS.svg", plot = last_plot(), dpi = 300, width = 8, height = 6, units = "in")
```


```{r}
# Check features in blanks and pools
data_blank <- data_final %>% dplyr::filter((str_detect(SampleID, regex("blank", ignore_case = TRUE)))) 
data_pools <- data_final %>% dplyr::filter((str_detect(SampleID, regex("Pool", ignore_case = TRUE))))
data_sixmix <- data_final %>% dplyr::filter(str_detect(SampleID, "^QC_P[0-9]+_[0-9]+$") | str_detect(SampleID, regex("6mix", ignore_case = TRUE)))
data_CTL <- data_final %>% dplyr::filter((str_detect(SampleID, regex("CTL", ignore_case = TRUE))))
```

```{r}
# Blanks
blanks_feature_info <- data.frame(Feature = colnames(data_blank)[-1],
                                  Mean_blank = data_blank %>% column_to_rownames("SampleID") %>% colMeans(), 
                                  SD_blank =  data_blank %>% column_to_rownames("SampleID") %>% apply(2, sd)) %>%
  dplyr::mutate(CV_blank = SD_blank/Mean_blank) %>% left_join(annotations, by = c("Feature" = "X.Scan.")) %>% 
  left_join(info_feature) %>% dplyr::select(Feature, mz, RT, Compound_Name, SpectrumID, 
                                            Precursor_MZ, Mean_blank, SD_blank, CV_blank) %>% 
  dplyr::filter(Mean_blank > 0) %>% arrange(desc(Mean_blank))

```
```{r}
# Six mix
sixmix_feature_info <- data.frame(Feature = colnames(data_sixmix)[-1],
                                  Mean_sixmix = data_sixmix %>% column_to_rownames("SampleID") %>% colMeans(), 
                                  SD_sixmix = data_sixmix %>% column_to_rownames("SampleID") %>% apply(2, sd)) %>%
  dplyr::mutate(CV_sixmix = SD_sixmix/Mean_sixmix) %>% left_join(annotations, by = c("Feature" = "X.Scan.")) %>% 
  left_join(info_feature) %>% dplyr::select(Feature, mz, RT, Compound_Name, SpectrumID, 
                                            Precursor_MZ, Mean_sixmix, SD_sixmix, CV_sixmix) %>% 
  dplyr::filter(Mean_sixmix > 0) %>% arrange(desc(Mean_sixmix))
```

```{r}
# Pool
pools_feature_info <- data.frame(Feature = colnames(data_pools)[-1],
                                 Mean_pool = data_pools %>% column_to_rownames("SampleID") %>% colMeans(), 
                                 SD_pool =  data_pools %>% column_to_rownames("SampleID") %>% apply(2, sd)) %>%
  dplyr::mutate(CV_pool = SD_pool/Mean_pool) %>% left_join(annotations, by = c("Feature" = "X.Scan.")) %>% 
  left_join(info_feature) %>% dplyr::select(Feature, mz, RT, Compound_Name, SpectrumID, 
                                            Precursor_MZ, Mean_pool, SD_pool, CV_pool) %>% 
  dplyr::filter(Mean_pool > 0) %>% arrange(desc(Mean_pool))

```
```{r}
# Features to be removed Pools/Blank < 5 or 10
feature_to_remove <- blanks_feature_info %>% left_join(pools_feature_info) %>%
  dplyr::filter(Mean_blank > 0) %>% 
  dplyr::mutate(Pool_Blank = Mean_pool/Mean_blank) %>% 
  dplyr::filter(Pool_Blank < 5 | is.na(Pool_Blank))
```

```{r}
# Data with blank removal
data_clean <- data_final %>% dplyr::select(-c(feature_to_remove$Feature)) %>%
  column_to_rownames("SampleID") %>%
  select_if(~sum(.) != 0) %>% rownames_to_column("SampleID")
```
```{r}
# Features to be removed Pool/QCmix < 5 or 10 
feature_to_remove_mix <- sixmix_feature_info %>% left_join(pools_feature_info) %>% 
  dplyr::filter(Mean_sixmix > 0) %>% 
  dplyr::mutate(Pool_Mix = Mean_pool/Mean_sixmix) %>% 
  dplyr::filter(Pool_Mix < 5 | is.na(Pool_Mix)) %>% dplyr::filter(!(Feature %in% feature_to_remove$Feature))

# Data with blank2 removal
data_clean2 <- data_clean %>% dplyr::select(-c(feature_to_remove_mix$Feature))

```
```{r}
# Define the color palette for six Sample_Type categories
sample_type_colors <- c("samples" = "#1f77b4",  # Blue
                        "Blank" = "#ff7f0e",  # Orange
                        "Pooled samples" = "#2ca02c",  # Green
                        "CTL" = "#d62728",  # Red
                        "6mix" = "#9467bd",  # Purple
                        "Wash" = "#8c564b")  # Brown
```

```{r}
# PCA raw data
PCA_raw <- mixOmics::pca(data_clean2 %>% select_at(vars(-one_of(nearZeroVar(., names = TRUE)))) %>%
                           column_to_rownames("SampleID"), ncomp = 2, center = TRUE, scale = TRUE)
PCA_raw_scores <- data.frame(PCA_raw$variates$X) %>% rownames_to_column("SampleID") %>% left_join(metadata_final)

i <- "Sample_Type"

PCA_raw_plot <- PCA_raw_scores %>%
  ggscatter(x = "PC1", y = "PC2", color = i, alpha = 0.6,
            title = paste("PCA - Skin Metabolome", i, sep = " "),
            xlab = paste("PC1 (", round(PCA_raw$prop_expl_var$X[1]*100, digits = 1),"%)", sep = ""), 
            ylab = paste("PC2 (", round(PCA_raw$prop_expl_var$X[2]*100, digits = 1),"%)", sep = ""),
            ggtheme = theme_classic()) +
  geom_point(data = PCA_raw_scores %>% group_by((!!sym(i))) %>% 
               summarise_at(vars(matches("PC")), mean), aes(PC1, PC2, color = (!!sym(i))), size = 4, shape = 8) +
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8),
        axis.text = element_text(size = 4)) + coord_fixed() +
  scale_color_manual(values = sample_type_colors)

```

```{r}
ggsave("../figures/PCA_qc_5.png", plot = last_plot(), dpi = 300, width = 10, height = 6, units = "in")
```

```{r}
ggsave("../figures/PCA_qc_5.svg", plot = last_plot(), dpi = 300, width = 10, height = 6, units = "in")
```
```{r}
data_clean_2_clr <- decostand(data_clean2 %>% column_to_rownames("SampleID"), method = "rclr")
```
```{r}
# PCA
PCA_whole <- mixOmics::pca(data_clean_2_clr, ncomp = 2, center = TRUE, scale = TRUE)
PCA_whole_scores <- data.frame(PCA_whole$variates$X) %>% 
  rownames_to_column("SampleID") %>% left_join(metadata_final)
PCA_whole_scores <- data.frame(PCA_whole$variates$X) %>% rownames_to_column("SampleID") %>% left_join(metadata_final)
i <- "Sample_Type"

PCA_plot <- PCA_whole_scores %>%
  ggscatter(x = "PC1", y = "PC2", color = i, alpha = 0.6,
            title = paste("PCA - Skin Metabolome - rclr transformed data"),
            xlab = paste("PC1 (", round(PCA_whole$prop_expl_var$X[1]*100, digits = 1),"%)", sep = ""), 
            ylab = paste("PC2 (", round(PCA_whole$prop_expl_var$X[2]*100, digits = 1),"%)", sep = ""),
            ggtheme = theme_classic()) +
  geom_point(data = PCA_whole_scores %>% group_by((!!sym(i))) %>% 
               summarise_at(vars(matches("PC")), mean), aes(PC1, PC2, color = (!!sym(i))), size = 4, shape = 8) +
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8),
        axis.text = element_text(size = 4)) + coord_fixed()  +
  scale_color_manual(values = sample_type_colors)
```
```{r}
ggsave("../figures/PCA_qc_rclr.svg", plot = last_plot(), dpi = 300, width = 10, height = 6, units = "in")
```



```{r}
f <- ggarrange(
  PCA_raw_plot, PCA_plot, 
  ncol = 2, nrow = 1, 
  labels = c("a", "b"),
  common.legend = TRUE,  # Use a common legend
  legend = "right"  , align ="hv"    # Place the legend at the bottom
)

```

```{r}
ggsave("../figures/PCA_combined.svg", plot = last_plot(), dpi = 300, width = 10, height = 6, units = "in")
```

```{r}
# Remove plastic contamination
feature_check <- info_feature %>% dplyr::select(1:3) %>% 
  dplyr::filter(Feature %in% colnames(data_clean)) %>% as_tibble()
colnames(feature_check) <- c("peak_id", "mz", "rt")
feature_check$mz <- as.double(feature_check$mz)
feature_check$rt <- as.double(feature_check$rt)
```


```{r}
# for 1 CTL sample 
feature_pool <- data_clean2 %>% 
  dplyr::filter(SampleID == "P4_A1_LAMI_CTL_19") %>% column_to_rownames("SampleID") %>%
  t() %>% as.data.frame() %>% rownames_to_column("peak_id")

feature_check_pool <- feature_check %>% left_join(feature_pool)

feature_check_pool$peak_id <- as.integer(feature_check_pool$peak_id)
colnames(feature_check_pool)[4] <- "intensity"
```


```{r}
# Remove PEGs
peak_table_PEG <- detectHomologues(feature_check_pool, mz_steps = c(44.02628),
                                   rt_min = 0.1, rt_max = 100, ppm_tolerance = 10, 
                                   min_series_length = 5, search_mode = "targeted", 
                                   step_mode = "increment", verbose = TRUE)

plotAnnotatedStatic(peak_table_PEG)
```
```{r}

sdb <- sdbCreate(peak_table_PEG, sample_origin = "data")


```
# do not run just in case you might be interested in some of those compounds that are still present
```{r}
# Remove untargeted 
peak_table_untargeted <- detectHomologues(feature_check_pool, mz_min = 10, mz_max = 50, 
                                          rt_min = 0.5, rt_max = 100, ppm_tolerance = 5, 
                                          min_series_length = 5, search_mode = "untargeted", 
                                          step_mode = "increment", verbose = TRUE)

plotAnnotatedStatic(peak_table_untargeted)
```
```{r}
#if running targeted and untargeted 
# Combine tables
overlapping_ids <- sdbCheckContained(sdb, peak_table_untargeted, ppm_tolerance = 0.5, rt_tolerance = 0.5)
sdb <- sdbPush(sdb, filter(peak_table_untargeted, !(homologue_id %in% overlapping_ids)))
```


```{r}
sdb$sample_peak_id <- as.character(sdb$sample_peak_id)
features_contaminat <- info_feature_complete %>% dplyr::filter(Feature %in% sdb$sample_peak_id)

data_clean3 <- data_clean2 %>% dplyr::select(-c(sdb$sample_peak_id))
```

```{r}
# Check one of the PEGs
peg <- data_final %>% dplyr::select(SampleID, `30108`) %>% left_join(metadata_final) 
colnames(peg)[2] <- "PEG"

peg %>% arrange(Order) %>% dplyr::mutate(LogPeg = log2(PEG+1)) %>%
  ggboxplot(x = "Sample_Type", y = "PEG", add = "jitter",
                  color = "Order", group = "Sample_Type", legend = "right") + scale_color_viridis_c()
```

```{r}
ggsave("../figures/sequence_order/PEGboxplot_30108.png", plot = last_plot(), dpi = 300, width = 11, height = 5, units = "in")
```
```{r}
ggsave("../figures/sequence_order/PEGboxplot_30108.svg", plot = last_plot(), dpi = 300, width = 11, height = 5, units = "in")
```
```{r}
ggsave("../figures/sequence_order/PEGboxplot_30108.svg", plot = last_plot(), dpi = 300, width = 11, height = 5, units = "in", device = cairo_svg)

```

```{r}
# PCA
PCA_raw <- mixOmics::pca(data_clean3 %>% select_at(vars(-one_of(nearZeroVar(., names = TRUE)))) %>%
                           column_to_rownames("SampleID"), ncomp = 2, center = TRUE, scale = TRUE)
PCA_raw_scores <- data.frame(PCA_raw$variates$X) %>% 
  rownames_to_column("SampleID") %>% left_join(metadata_final)

i <- "Sample_Type"

PCA_raw_plot <- PCA_raw_scores %>%
  ggscatter(x = "PC1", y = "PC2", color = i, alpha = 0.6,
            title = paste("PCA - Skin Metabolome", i, sep = " "),
            xlab = paste("PC1 (", round(PCA_raw$prop_expl_var$X[1]*100, digits = 1),"%)", sep = ""), 
            ylab = paste("PC2 (", round(PCA_raw$prop_expl_var$X[2]*100, digits = 1),"%)", sep = ""),
            ggtheme = theme_classic()) +
  geom_point(data = PCA_raw_scores %>% group_by((!!sym(i))) %>% 
               summarise_at(vars(matches("PC")), mean), aes(PC1, PC2, color = (!!sym(i))), size = 4, shape = 8) +
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8),
        axis.text = element_text(size = 4)) + coord_fixed()

```

```{r}
ggsave("../figures/PCA_raw_HD.png", plot = last_plot(), dpi = 300, width = 8, height = 6, units = "in")
```

```{r}
# Keep only samples
data_sample <- data_clean3 %>% 
   dplyr::filter(!(str_detect(SampleID, regex("Blank|6mix|Pool|wash|CTL|QC", ignore_case = TRUE)))) %>% 
  column_to_rownames("SampleID") %>% 
  select_at(vars(-one_of(nearZeroVar(., names = TRUE)))) %>% rownames_to_column("SampleID")
```
```{r}
#write_csv(x = data_sample, file = "../output/data_sample_09272024.csv")

write_csv(x = data_sample, file = "../output/data_sample_10022024.csv")
```

```{r}
# Assuming metadata_final contains SampleID and sample_name columns

# Replace SampleID with sample_name
data_sample_name <- data_clean3 %>% 
  dplyr::filter(!(str_detect(SampleID, regex("Blank|6mix|Pool|wash|CTL|QC", ignore_case = TRUE)))) %>%
  dplyr::left_join(metadata_final %>% select(SampleID, sample_name), by = "SampleID") %>%
  dplyr::mutate(SampleID = sample_name) %>%
  dplyr::select(-sample_name) %>%
  column_to_rownames("SampleID") %>% 
  select_at(vars(-one_of(nearZeroVar(., names = TRUE)))) %>% 
  rownames_to_column("SampleID")

```

```{r}
#write_csv(x = data_sample, file = "../output/data_sample_09272024.csv")

write_csv(x = data_sample_name, file = "../output/data_sample_10022024.csv")
```

```{r}
# Replace SampleID with sample_name 09 29 2023
data_sample_name <- data_sample %>% 
  dplyr::filter(!(str_detect(SampleID, regex("Blank|6mix|Pool|wash|CTL|QC", ignore_case = TRUE)))) %>%
  dplyr::left_join(metadata_final %>% select(SampleID, sample_name), by = "SampleID") %>%
  dplyr::mutate(SampleID = sample_name) %>%
  dplyr::select(-sample_name) %>%
  column_to_rownames("SampleID") %>% 
  select_at(vars(-one_of(nearZeroVar(., names = TRUE)))) %>% 
  rownames_to_column("SampleID")
```

```{r}
write_csv(x = data_sample_name, file = "../output/quantwithoutblancs_09262024.csv")
```
```{r}
write_csv(x = data_sample_name, file = "../output/data_sample_09292024.csv")
```

```{r}
# Assuming 'data_sample' contains a column 'SampleID' and 'metadata_final' has 'SampleID' and 'Plate_Number'
# Step 1: Get the SampleIDs that have Plate_Number equal to 2
exclude_samples <- metadata_final %>%
  filter(Plate_Number == 2) %>%
  pull(SampleID)

# Step 2: Remove those SampleIDs from data_sample
data_sample_filtered <- data_sample %>%
  filter(!SampleID %in% exclude_samples)

# Now, 'data_sample_filtered' contains only the rows where SampleID is not from Plate 2
```
```{r}
# Filter out rows where Plate_Number is equal to 2
metadata_filtered <- metadata_final %>%
  filter(Plate_Number != 2)
```
```{r}
# Replace SampleID with sample_name
data_sample_name <- data_sample_filtered %>% 
  dplyr::filter(!(str_detect(SampleID, regex("Blank|6mix|Pool|wash|CTL|QC", ignore_case = TRUE)))) %>%
  dplyr::left_join(metadata_filtered %>% select(SampleID, sample_name), by = "SampleID") %>%
  dplyr::mutate(SampleID = sample_name) %>%
  dplyr::select(-sample_name) %>%
  column_to_rownames("SampleID") %>% 
  select_at(vars(-one_of(nearZeroVar(., names = TRUE)))) %>% 
  rownames_to_column("SampleID")
```


```{r}
write_csv(x = metadata_filtered, file = "../output/metadata_filtered_09292024.csv")
```
```{r}
write_csv(x = metadata_final, file = "../output/metadata_09262024.csv")
```

```{r}
write_csv(x = data_sample_name, file = "../output/data_sample_filtered_09292024.csv")
```

```{r}
data_sample <- data_sample_filtered
```

```{r}
# RCLR transformation
data_sample_clr <- decostand(data_sample %>% column_to_rownames("SampleID"), method = "rclr")
```


```{r}
# PCA
PCA_whole <- mixOmics::pca(data_sample_clr, ncomp = 2, center = TRUE, scale = TRUE)
PCA_whole_scores <- data.frame(PCA_whole$variates$X) %>% 
  rownames_to_column("SampleID") %>% left_join(metadata_final)
PCA_whole_scores$group <- factor(PCA_whole_scores$group, levels = c("Healthy", "Acne_L", "Acne_NL"))

i <- "group"

PCA_plot <- PCA_whole_scores %>%
  ggscatter(x = "PC1", y = "PC2", color = i, alpha = 0.6,
            title = paste("PCA - group"),
            xlab = paste("PC1 (", round(PCA_whole$prop_expl_var$X[1]*100, digits = 1),"%)", sep = ""), 
            ylab = paste("PC2 (", round(PCA_whole$prop_expl_var$X[2]*100, digits = 1),"%)", sep = ""),
            ggtheme = theme_classic()) +
  geom_point(data = PCA_whole_scores %>% group_by((!!sym(i))) %>% 
               summarise_at(vars(matches("PC")), mean), aes(PC1, PC2, color = (!!sym(i))), size = 4, shape = 8) +
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8),
        axis.text = element_text(size = 4)) + coord_fixed() + scale_color_viridis_d()

```
```{r}
ggsave("../figures/PCA_plot_group.png", plot = last_plot(), dpi = 300, width = 8, height = 5, units = "in")
```
```{r}
#Excluding means
i <- "subject_ID"
j <- "group"

PCA_plot_2 <- PCA_whole_scores %>%
  ggscatter(x = "PC1", y = "PC2", color = i, shape = j, alpha = 0.6,
            title = paste("PCA - Skin Metabolic Profiles Acne lesional and non-lesional vs Healthy "),
            xlab = paste("PC1 (", round(PCA_whole$prop_expl_var$X[1]*100, digits = 1),"%)", sep = ""), 
            ylab = paste("PC2 (", round(PCA_whole$prop_expl_var$X[2]*100, digits = 1),"%)", sep = ""),
            ggtheme = theme_classic()) +
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8),
        axis.text = element_text(size = 4)) + coord_fixed()

```
```{r}
ggsave("../figures/PCA_plot_group_subjectID.png", plot = last_plot(), dpi = 300, width = 8, height = 8, units = "in")
```

```{r}
#including means 
i <- "subject_ID"
j <- "group"

PCA_plot_2 <- PCA_whole_scores %>%
  ggscatter(x = "PC1", y = "PC2", color = i, shape = j, alpha = 0.6,
            title = paste("PCA - Skin Metabolic Profiles Acne lesional and non-lesional vs Healthy "),
            xlab = paste("PC1 (", round(PCA_whole$prop_expl_var$X[1]*100, digits = 1),"%)", sep = ""), 
            ylab = paste("PC2 (", round(PCA_whole$prop_expl_var$X[2]*100, digits = 1),"%)", sep = ""),
            ggtheme = theme_classic()) +
  geom_point(data = PCA_whole_scores %>% group_by((!!sym(i))) %>% 
               summarise_at(vars(matches("PC")), mean), aes(PC1, PC2, color = (!!sym(i))), size = 4, shape = 8) + 
  # Remove color aesthetic from the grouping mean points
  geom_point(data = PCA_whole_scores %>% group_by((!!sym(j))) %>% 
               summarise_at(vars(matches("PC")), mean), aes(PC1, PC2), size = 4, shape = 8) +  # Shape only
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8),
        axis.text = element_text(size = 4)) + coord_fixed()

```

```{r}
# PERMANOVA
dist_metabolites <- vegdist(data_sample_clr, method = "euclidean")
disper_donor <- betadisper(dist_metabolites, PCA_whole_scores$subject_ID)

#significant variation between the groups
anova(disper_donor)
permanova <- adonis2(dist_metabolites ~ PCA_whole_scores$subject_ID + PCA_whole_scores$group, PCA_whole_scores, na.action = na.omit)

```


#ACNE_L vs Healthy
```{r}
# Acne_L vs Healthy
sample_LvsH <- metadata_final %>% 
  dplyr::filter(group == "Acne_L" | group == "Healthy")

data_LvsH <- data_sample %>% 
  dplyr::filter(SampleID %in% sample_LvsH$SampleID) %>% 
  column_to_rownames("SampleID") %>% 
  select_at(vars(-one_of(nearZeroVar(., names = TRUE)))) %>% 
  rownames_to_column("SampleID")
```

```{r}
# RCLR transformation
data_LvsH_clr <- decostand(data_LvsH %>% column_to_rownames("SampleID"), method = "rclr")

```


```{r}
# PCA
PCA_LvsH <- mixOmics::pca(data_LvsH_clr, ncomp = 2, center = TRUE, scale = TRUE)
PCA_LvsH_scores <- data.frame(PCA_LvsH$variates$X) %>% 
  rownames_to_column("SampleID") %>% left_join(metadata_final)
```

```{r}
i <- "group"

PCA_LvsH_plot <- PCA_LvsH_scores %>%
  ggscatter(x = "PC1", y = "PC2", color = i, alpha = 0.6,
            title = paste("PCA - Acne Lesional vs Healthy "), palette = "npg",
            xlab = paste("PC1 (", round(PCA_LvsH$prop_expl_var$X[1]*100, digits = 1),"%)", sep = ""), 
            ylab = paste("PC2 (", round(PCA_LvsH$prop_expl_var$X[2]*100, digits = 1),"%)", sep = ""),
            ggtheme = theme_classic()) +
  geom_point(data = PCA_LvsH_scores %>% group_by((!!sym(i))) %>% 
               summarise_at(vars(matches("PC")), mean), aes(PC1, PC2, color = (!!sym(i))), size = 4, shape = 8) +
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8),
        axis.text = element_text(size = 4)) + coord_fixed()
```

```{r}
i <- "group"
#"#a54700", "#008080"

PCA_LvsH_plot <- PCA_LvsH_scores %>%
  ggscatter(x = "PC1", y = "PC2", color = i, alpha = 0.6,
            title = paste("PCA - Acne Lesional vs Healthy "), palette = c("#dd7966", "#58a35b"),
            xlab = paste("PC1 (", round(PCA_LvsH$prop_expl_var$X[1]*100, digits = 1),"%)", sep = ""), 
            ylab = paste("PC2 (", round(PCA_LvsH$prop_expl_var$X[2]*100, digits = 1),"%)", sep = ""),
            ggtheme = theme_classic()) +
  geom_point(data = PCA_LvsH_scores %>% group_by(group) %>% 
               summarise_at(vars(matches("PC")), mean), aes(PC1, PC2, color = group), size = 4, shape = 8) +
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8),
        axis.text = element_text(size = 4)) + coord_fixed()
```

```{r}
ggsave("../figures/PCA_plot_Acne_LvsHealthy.png", plot = last_plot(), dpi = 300, width = 8, height = 6, units = "in")
```

```{r}
#Permanova 
dist_metabolites <- vegdist(data_LvsH_clr, method = "euclidean")
disper_group <- betadisper(dist_metabolites, PCA_LvsH_scores$group)
anova(disper_group)
permanova <- adonis2(dist_metabolites ~ PCA_LvsH_scores$subject_ID + PCA_LvsH_scores$group, PCA_LvsH_scores, na.action = na.omit)

permanova <- adonis2(dist_metabolites ~ PCA_LvsH_scores$subject_ID + PCA_LvsH_scores$group, PCA_LvsH_scores, na.action = na.omit)
```

```{r}
permanova <- adonis2(dist_metabolites ~ PCA_LvsH_scores$subject_ID + PCA_LvsH_scores$group, 
                     data = PCA_LvsH_scores, 
                     na.action = na.omit, 
                     by = "margin")

```

#The subject-specific differences (likely individual biological variability) are driving the majority of the variation in metabolite distances.
#The grouping factor (Lesional vs Healthy) doesn't explain additional variation when individual differences are already considered.
```{r}
# PLSDA
PLSDA_LvsH <- mixOmics::plsda(data_LvsH_clr, PCA_LvsH_scores$group, ncomp = 2, scale = TRUE)
PLSDA_LvsH_scores <- data.frame(PLSDA_LvsH$variates$X) %>% 
  rownames_to_column("SampleID") %>% left_join(metadata_final)

PLSDA_LvsH_plot <- PLSDA_LvsH_scores %>%
  ggscatter(x = "comp1", y = "comp2", color = "group", alpha = 0.6, title = "PLS-DA - Acne Lesional vs Healthy", 
            xlab = paste("Component 1 (", round(PLSDA_LvsH$prop_expl_var$X[1]*100, digits = 1),"%)", sep = ""), 
            ylab = paste("Component 2 (", round(PLSDA_LvsH$prop_expl_var$X[2]*100, digits = 1),"%)", sep = ""),
            legend.title = "Group", ggtheme = theme_classic(), palette = c("#dd7966", "#58a35b")) +
  geom_point(data = PLSDA_LvsH_scores %>% group_by(group) %>% 
               summarise_at(vars(matches("comp")), mean), aes(comp1, comp2, color = group), size = 3, shape = 8) +
  theme(plot.title = element_text(size = 10),axis.title = element_text(size = 8),
        axis.text = element_text(size = 5)) + coord_fixed()
```


#palette = c("#F8766D", "#619CFF"))

```{r}
ggsave("../figures/PLSDA_plot_Acne_LvsHealthy.png", plot = last_plot(), dpi = 300, width = 8, height = 6, units = "in")
```
```{r}
Loadings_LvsH <- plotLoadings(PLSDA_LvsH, contrib = "max") %>%
  as.data.frame(check.names = FALSE) %>%
  rownames_to_column()
```



```{r}
# Dynamically remove 'X.' prefix if it exists
colnames(Loadings_LvsH) <- gsub("^X\\.", "", colnames(Loadings_LvsH))

# Now, select the columns of interest
Loadings_LvsH <- Loadings_LvsH %>%
  dplyr::select(rowname, GroupContrib, importance)

perf_PLSDA_LvsH <- perf(PLSDA_LvsH, validation = "Mfold", folds =5) #perform M-fold or loo
plot(perf_PLSDA_LvsH, legend = FALSE)

VIPs_LvsH <- as.data.frame(mixOmics::vip(PLSDA_LvsH))
VIPs_LvsH_filter <- dplyr::filter(VIPs_LvsH, VIPs_LvsH$comp1 > 1) # larger than 1 is seen as an important component
VIPs_LvsH_filter$ID <- rownames(VIPs_LvsH_filter)
VIPs_LvsH_select <- VIPs_LvsH_filter %>% dplyr::select(ID, comp1)
VIPs_LvsH_Load <- VIPs_LvsH_select %>% 
  left_join(Loadings_LvsH, by = c("ID" = "rowname")) %>% arrange(desc(comp1)) %>%
  left_join(info_feature_complete, by = c("ID" = "Feature")) %>% dplyr::select(1:8)
```


```{r}
write_csv(x = VIPs_LvsH_Load, file = "../output/VIPs_LvsH_Load.csv")
```

```{r}
skin_Acne_L <- VIPs_LvsH_Load %>% dplyr::filter(GroupContrib == "Acne_L") 
skin_Healthy <- VIPs_LvsH_Load %>% dplyr::filter(GroupContrib == "Healthy") 
```

```{r}
data_hfhc <- data_sample %>%
  dplyr::select("SampleID", skin_Acne_L$ID, skin_Healthy$ID) %>%
  dplyr::mutate(Healthy = rowSums(select(., skin_Healthy$ID))) %>%
  dplyr::mutate(Acne_L = rowSums(select(., skin_Acne_L$ID))) %>%
  dplyr::mutate(Ratio = log(Acne_L/Healthy)) %>%
  left_join(metadata_final) %>%
  dplyr::mutate(group = factor(group, levels = c("Healthy", "Acne_L", "Acne_NL")))

```

```{r}
# Calculate the mean per group and age
df_mean <- data_hfhc %>%
  group_by(group, local_lesion_severity) %>%
  summarise(Mean_Measurement = mean(Ratio))
```


```{r}
# "#008080", "#a54700", "#c49c6b"
# Plot
ratio_hfhc_plot <- ggplot(data_hfhc, aes(x = local_lesion_severity, y = Ratio,)) +
  geom_line(aes(alpha = 0.3, colour = group, group = subject_ID), linewidth = 0.3) +  # Thin lines for individual mice
  geom_line(data = df_mean, aes(x = local_lesion_severity, y = Mean_Measurement, color = group), 
            linewidth = 2.5) +  # Thick lines for the mean
  scale_color_manual(values = c("#dd7966","#6ab2bd", "#58a35b")) +  # Customize colors if needed
  theme_bw() +
  labs(x = "Local Lesion Severity", y = "Ratio") +
  theme(legend.position = "none")
```



```{r}
data_hfhc_plot <- data_hfhc %>% 
  ggboxplot(x = "group", y = "Ratio", legend = "none",
            title = "Driving Metabolic Features", 
            ylab = "Ratio (Top Acne_L/Top Healthy)", xlab = FALSE,
            add = "jitter", palette = c("#58a35b", "#dd7966", "#6ab2bd"),
            add.params = list(color = "group", alpha = 0.6)) + 
  stat_compare_means(comparisons = list(c("Acne_L", "Healthy"), c("Acne_L", "Acne_NL"),
                                        c("Healthy", "Acne_NL")), label = "p.signif") + 
  theme(plot.title = element_text(size = 10), 
        axis.title.y = element_text(size = 10), 
        axis.text.y = element_text(size = 8), 
        axis.text.x = element_text(size = 8), # Ensure x-axis text is visible and not blank
        axis.ticks.x = element_line())        # Ensure x-axis ticks are visible
anova <- aov(Ratio ~ group, data_hfhc)
tukey_hsd(anova)
```
```{r}
ggsave("../figures/boxplot_ratioAcne_LvsHealthy.png", plot = last_plot(), dpi = 300, width = 6, height = 4, units = "in")
```

write_csv(x = metadata_final, file = "../output/metadata_09262024.csv")

#ACNE_NL vs Healthy
```{r}
# Acne_NL vs Healthy
sample_NLvsH <- metadata_final %>% 
  dplyr::filter(group == "Acne_NL" | group == "Healthy")

data_NLvsH <- data_sample %>% 
  dplyr::filter(SampleID %in% sample_NLvsH$SampleID) %>% 
  column_to_rownames("SampleID") %>% 
  select_at(vars(-one_of(nearZeroVar(., names = TRUE)))) %>% 
  rownames_to_column("SampleID")
```

```{r}
# RCLR transformation
data_NLvsH_clr <- decostand(data_NLvsH %>% column_to_rownames("SampleID"), method = "rclr")

```


```{r}
# PCA
PCA_NLvsH <- mixOmics::pca(data_NLvsH_clr, ncomp = 2, center = TRUE, scale = TRUE)
PCA_NLvsH_scores <- data.frame(PCA_NLvsH$variates$X) %>% 
  rownames_to_column("SampleID") %>% left_join(metadata_final)
```

```{r}
i <- "group"

PCA_NLvsH_plot <- PCA_NLvsH_scores %>%
  ggscatter(x = "PC1", y = "PC2", color = i, alpha = 0.6,
            title = paste("PCA - Acne Non-Lesional vs Healthy "), palette = "npg",
            xlab = paste("PC1 (", round(PCA_NLvsH$prop_expl_var$X[1]*100, digits = 1),"%)", sep = ""), 
            ylab = paste("PC2 (", round(PCA_NLvsH$prop_expl_var$X[2]*100, digits = 1),"%)", sep = ""),
            ggtheme = theme_classic()) +
  geom_point(data = PCA_NLvsH_scores %>% group_by((!!sym(i))) %>% 
               summarise_at(vars(matches("PC")), mean), aes(PC1, PC2, color = (!!sym(i))), size = 4, shape = 8) +
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8),
        axis.text = element_text(size = 4)) + coord_fixed()
```

```{r}
i <- "group"
#"#c49c6b", "#008080"

PCA_NLvsH_plot <- PCA_NLvsH_scores %>%
  ggscatter(x = "PC1", y = "PC2", color = i, alpha = 0.6,
            title = paste("PCA - Acne Non-Lesional vs Healthy "), palette = c("#6ab2bd", "#58a35b"),
            xlab = paste("PC1 (", round(PCA_NLvsH$prop_expl_var$X[1]*100, digits = 1),"%)", sep = ""), 
            ylab = paste("PC2 (", round(PCA_NLvsH$prop_expl_var$X[2]*100, digits = 1),"%)", sep = ""),
            ggtheme = theme_classic()) +
  geom_point(data = PCA_NLvsH_scores %>% group_by(group) %>% 
               summarise_at(vars(matches("PC")), mean), aes(PC1, PC2, color = group), size = 4, shape = 8) +
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8),
        axis.text = element_text(size = 4)) + coord_fixed()
```

```{r}
ggsave("../figures/PCA_plot_Acne_NLvsHealthy.png", plot = last_plot(), dpi = 300, width = 8, height = 6, units = "in")
```

```{r}
#Permanova 
dist_metabolites <- vegdist(data_NLvsH_clr, method = "euclidean")
disper_group <- betadisper(dist_metabolites, PCA_NLvsH_scores$group)
anova(disper_group)
permanova <- adonis2(dist_metabolites ~ PCA_NLvsH_scores$subject_ID + PCA_NLvsH_scores$group, PCA_NLvsH_scores, na.action = na.omit)

```
```{r}
permanova <- adonis2(dist_metabolites ~ PCA_NLvsH_scores$subject_ID + PCA_NLvsH_scores$group, 
                     data = PCA_NLvsH_scores, 
                     na.action = na.omit, 
                     by = "margin")
```

```{r}
# PLSDA
#"#c49c6b", "#008080"
PLSDA_NLvsH <- mixOmics::plsda(data_NLvsH_clr, PCA_NLvsH_scores$group, ncomp = 2, scale = TRUE)
PLSDA_NLvsH_scores <- data.frame(PLSDA_NLvsH$variates$X) %>% 
  rownames_to_column("SampleID") %>% left_join(metadata_final)

PLSDA_NLvsH_plot <- PLSDA_NLvsH_scores %>%
  ggscatter(x = "comp1", y = "comp2", color = "group", alpha = 0.6, title = "PLS-DA - Acne Non-Lesional vs Healthy", 
            xlab = paste("Component 1 (", round(PLSDA_NLvsH$prop_expl_var$X[1]*100, digits = 1),"%)", sep = ""), 
            ylab = paste("Component 2 (", round(PLSDA_NLvsH$prop_expl_var$X[2]*100, digits = 1),"%)", sep = ""),
            legend.title = "Group", ggtheme = theme_classic(), palette = c("#6ab2bd", "#58a35b")) +
  geom_point(data = PLSDA_NLvsH_scores %>% group_by(group) %>% 
               summarise_at(vars(matches("comp")), mean), aes(comp1, comp2, color = group), size = 3, shape = 8) +
  theme(plot.title = element_text(size = 10),axis.title = element_text(size = 8),
        axis.text = element_text(size = 5)) + coord_fixed()
```


#palette = c("#F8766D", "#619CFF"))

```{r}
ggsave("../figures/PLSDA_plot_Acne_NLvsHealthy.png", plot = last_plot(), dpi = 300, width = 8, height = 6, units = "in")
```
```{r}

Loadings_NLvsH <- plotLoadings(PLSDA_NLvsH, contrib = "max") %>%
  as.data.frame(check.names = FALSE) %>%
  rownames_to_column()

# Dynamically remove 'X.' prefix if it exists
colnames(Loadings_NLvsH) <- gsub("^X\\.", "", colnames(Loadings_NLvsH))

# Now, select the columns of interest
Loadings_NLvsH <- Loadings_NLvsH %>%
  dplyr::select(rowname, GroupContrib, importance)

perf_PLSDA_NLvsH <- perf(PLSDA_NLvsH, validation = "Mfold", folds = 5 ) #perform M-fold or loo
plot(perf_PLSDA_NLvsH, legend = FALSE)

VIPs_NLvsH <- as.data.frame(mixOmics::vip(PLSDA_NLvsH))
VIPs_NLvsH_filter <- dplyr::filter(VIPs_NLvsH, VIPs_NLvsH$comp1 > 1) # larger than 1 is seen as an important component
VIPs_NLvsH_filter$ID <- rownames(VIPs_NLvsH_filter)
VIPs_NLvsH_select <- VIPs_NLvsH_filter %>% dplyr::select(ID, comp1)
VIPs_NLvsH_Load <- VIPs_NLvsH_select %>% 
  left_join(Loadings_NLvsH, by = c("ID" = "rowname")) %>% arrange(desc(comp1)) %>%
  left_join(info_feature_complete, by = c("ID" = "Feature")) %>% dplyr::select(1:8)
```


```{r}
write_csv(x = VIPs_NLvsH_Load, file = "../output/VIPs_NLvsH_Load.csv")
```

#ACNE_NL vs Acne_L
```{r}
# Acne_NL vs Acne_L
sample_NLvsL <- metadata_final %>% 
  dplyr::filter(group == "Acne_NL" | group == "Acne_L")

data_NLvsL <- data_sample %>% 
  dplyr::filter(SampleID %in% sample_NLvsL$SampleID) %>% 
  column_to_rownames("SampleID") %>% 
  select_at(vars(-one_of(nearZeroVar(., names = TRUE)))) %>% 
  rownames_to_column("SampleID")
```

```{r}
# RCLR transformation
data_NLvsL_clr <- decostand(data_NLvsL %>% column_to_rownames("SampleID"), method = "rclr")

```


```{r}
# PCA
PCA_NLvsL <- mixOmics::pca(data_NLvsL_clr, ncomp = 2, center = TRUE, scale = TRUE)
PCA_NLvsL_scores <- data.frame(PCA_NLvsL$variates$X) %>% 
  rownames_to_column("SampleID") %>% left_join(metadata_final)
```

```{r}
i <- "group"

PCA_NLvsL_plot <- PCA_NLvsL_scores %>%
  ggscatter(x = "PC1", y = "PC2", color = i, alpha = 0.6,
            title = paste("PCA - Acne Non-Lesional vs Lesional "), palette = "npg",
            xlab = paste("PC1 (", round(PCA_NLvsL$prop_expl_var$X[1]*100, digits = 1),"%)", sep = ""), 
            ylab = paste("PC2 (", round(PCA_NLvsL$prop_expl_var$X[2]*100, digits = 1),"%)", sep = ""),
            ggtheme = theme_classic()) +
  geom_point(data = PCA_NLvsL_scores %>% group_by((!!sym(i))) %>% 
               summarise_at(vars(matches("PC")), mean), aes(PC1, PC2, color = (!!sym(i))), size = 4, shape = 8) +
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8),
        axis.text = element_text(size = 4)) + coord_fixed()
```

```{r}
i <- "group"
#"#58a35b", "#dd7966", "#6ab2bd"
#"#a54700","#c49c6b"
PCA_NLvsL_plot <- PCA_NLvsL_scores %>%
  ggscatter(x = "PC1", y = "PC2", color = i, alpha = 0.6,
            title = paste("PCA - Acne Non-Lesional vs Lesional "), palette = c("#dd7966","#6ab2bd"),
            xlab = paste("PC1 (", round(PCA_NLvsL$prop_expl_var$X[1]*100, digits = 1),"%)", sep = ""), 
            ylab = paste("PC2 (", round(PCA_NLvsL$prop_expl_var$X[2]*100, digits = 1),"%)", sep = ""),
            ggtheme = theme_classic()) +
  geom_point(data = PCA_NLvsL_scores %>% group_by(group) %>% 
               summarise_at(vars(matches("PC")), mean), aes(PC1, PC2, color = group), size = 4, shape = 8) +
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8),
        axis.text = element_text(size = 4)) + coord_fixed()
```

```{r}
ggsave("../figures/PCA_plot_Acne_NLvsL.png", plot = last_plot(), dpi = 300, width = 8, height = 6, units = "in")
```

```{r}
#Permanova 
dist_metabolites <- vegdist(data_NLvsL_clr, method = "euclidean")
disper_group <- betadisper(dist_metabolites, PCA_NLvsL_scores$group)
anova(disper_group)
permanova <- adonis2(dist_metabolites ~ PCA_NLvsL_scores$subject_ID + PCA_NLvsL_scores$group, PCA_NLvsL_scores, na.action = na.omit)

```
```{r}
permanova <- adonis2(dist_metabolites ~ PCA_NLvsL_scores$subject_ID + PCA_NLvsL_scores$group, 
                     data = PCA_NLvsL_scores, 
                     na.action = na.omit, 
                     by = "margin")
```

```{r}
# PLSDA
PLSDA_NLvsL <- mixOmics::plsda(data_NLvsL_clr, PCA_NLvsL_scores$group, ncomp = 2, scale = TRUE)
PLSDA_NLvsL_scores <- data.frame(PLSDA_NLvsL$variates$X) %>% 
  rownames_to_column("SampleID") %>% left_join(metadata_final)

PLSDA_NLvsL_plot <- PLSDA_NLvsL_scores %>%
  ggscatter(x = "comp1", y = "comp2", color = "group", alpha = 0.6, title = "PLS-DA - Acne Non-Lesional vs Lesional", 
            xlab = paste("Component 1 (", round(PLSDA_NLvsL$prop_expl_var$X[1]*100, digits = 1),"%)", sep = ""), 
            ylab = paste("Component 2 (", round(PLSDA_NLvsL$prop_expl_var$X[2]*100, digits = 1),"%)", sep = ""),
            legend.title = "Group", ggtheme = theme_classic(), palette = c("#dd7966","#6ab2bd")) +
  geom_point(data = PLSDA_NLvsL_scores %>% group_by(group) %>% 
               summarise_at(vars(matches("comp")), mean), aes(comp1, comp2, color = group), size = 3, shape = 8) +
  theme(plot.title = element_text(size = 10),axis.title = element_text(size = 8),
        axis.text = element_text(size = 5)) + coord_fixed()
```


#palette = c("#F8766D", "#619CFF"))

```{r}
ggsave("../figures/PLSDA_plot_Acne_NLvsL.png", plot = last_plot(), dpi = 300, width = 8, height = 6, units = "in")
```
```{r}

Loadings_NLvsL <- plotLoadings(PLSDA_NLvsL, contrib = "max") %>%
  as.data.frame(check.names = FALSE) %>%
  rownames_to_column()

# Dynamically remove 'X.' prefix if it exists
colnames(Loadings_NLvsL) <- gsub("^X\\.", "", colnames(Loadings_NLvsL))

# Now, select the columns of interest
Loadings_NLvsL <- Loadings_NLvsL %>%
  dplyr::select(rowname, GroupContrib, importance)

perf_PLSDA_NLvsL <- perf(PLSDA_NLvsL, validation = "Mfold", folds=5) #perform M-fold or loo
plot(perf_PLSDA_NLvsL, legend = FALSE)

VIPs_NLvsL <- as.data.frame(mixOmics::vip(PLSDA_NLvsL))
VIPs_NLvsL_filter <- dplyr::filter(VIPs_NLvsL, VIPs_NLvsL$comp1 > 1) # larger than 1 is seen as an important component
VIPs_NLvsL_filter$ID <- rownames(VIPs_NLvsL_filter)
VIPs_NLvsL_select <- VIPs_NLvsL_filter %>% dplyr::select(ID, comp1)
VIPs_NLvsL_Load <- VIPs_NLvsL_select %>% 
  left_join(Loadings_NLvsL, by = c("ID" = "rowname")) %>% arrange(desc(comp1)) %>%
  left_join(info_feature_complete, by = c("ID" = "Feature")) %>% dplyr::select(1:8)
```


```{r}
combined_pca <- ggarrange(PCA_LvsH_plot, PCA_NLvsH_plot, PCA_NLvsL_plot, PLSDA_LvsH_plot, PLSDA_NLvsH_plot, PLSDA_NLvsL_plot, ncol = 3, nrow =2, labels = c("a", "b", "c", "d", "e", "f"), align ="hv", common.legend = FALSE)

```

```{r}
ggsave("../figures/combined_PCA_PLSDA.png", plot = last_plot(), dpi = 300, width = 14, height = 7, units = "in")
```

#######################################################
# Feature affected by Acne and exacerbated by Lesions #
#######################################################

```{r}
#comparison included in first cytoscape
feature_interest <- VIPs_LvsH_Load %>% dplyr::select(ID, comp1, GroupContrib) %>%
  dplyr::rename(vip_LvsH = "comp1", group_LvsH = "GroupContrib") %>% 
  full_join(VIPs_NLvsH_Load) %>% 
  dplyr::rename(vip_NLvsH = "comp1", group_NLvsH = "GroupContrib") %>%
  full_join(VIPs_NLvsL_Load) %>% 
  dplyr::rename(vip_NLvsL = "comp1", group_NLvsL = "GroupContrib") %>%
  dplyr::mutate(combined_vip = rowSums(cbind(vip_LvsH, vip_NLvsH, vip_NLvsL), na.rm = TRUE)) %>%
  dplyr::mutate(Group = paste(group_LvsH, group_NLvsH, group_NLvsL, sep = "_")) %>%
  dplyr::mutate(Outcome = case_when(group_LvsH == "Acne_L" & group_NLvsH == "Acne_NL" & is.na(group_NLvsL) ~ "Increased in Acne",
                                    group_LvsH == "Healthy" & group_NLvsH == "Healthy" & is.na(group_NLvsL) ~ "Decreased in Acne",
                                    group_LvsH == "Acne_L" & is.na(group_NLvsH) & group_NLvsL == "Acne_L" ~ "Increased in Acne Lesions",
                                    is.na(group_LvsH) & is.na(group_NLvsH) & group_NLvsL == "Acne_L" ~ "Increased in Acne Lesions compared to Non-Lesions",
                                    group_LvsH=="Acne_L" & is.na(group_NLvsH) & is.na(group_NLvsL) ~ "Increased in Acne Lesions compared to Healthy",
                                    TRUE ~ "remove")) %>% dplyr::filter(Outcome != "remove")
    
```
```{r}
feature_interest <- VIPs_LvsH_Load %>% dplyr::select(ID, comp1, GroupContrib) %>%
  dplyr::rename(vip_LvsH = "comp1", group_LvsH = "GroupContrib") %>% 
  full_join(VIPs_NLvsH_Load) %>% 
  dplyr::rename(vip_NLvsH = "comp1", group_NLvsH = "GroupContrib") %>%
  full_join(VIPs_NLvsL_Load) %>% 
  dplyr::rename(vip_NLvsL = "comp1", group_NLvsL = "GroupContrib") %>%
  dplyr::mutate(combined_vip = rowSums(cbind(vip_LvsH, vip_NLvsH, vip_NLvsL), na.rm = TRUE)) %>%
  dplyr::mutate(Group = paste(group_LvsH, group_NLvsH, group_NLvsL, sep = "_")) %>%
  dplyr::mutate(Outcome = case_when(group_LvsH == "Acne_L" & group_NLvsH == "Acne_NL" & is.na(group_NLvsL) ~ "Increased in Acne",
                                    group_LvsH == "Healthy" & group_NLvsH == "Healthy" & is.na(group_NLvsL) ~ "Decreased in Acne",
                                    group_LvsH == "Acne_L" & is.na(group_NLvsH) & group_NLvsL == "Acne_L" ~ "Increased in Acne Lesions",
                                    is.na(group_LvsH) & is.na(group_NLvsH) & group_NLvsL == "Acne_L" ~ "Increased in Acne Lesions compared to Non-Lesions",
                                    group_LvsH=="Acne_L" & is.na(group_NLvsH) & is.na(group_NLvsL) ~ "Increased in Acne Lesions compared to Healthy",
                                    is.na(group_LvsH) & group_NLvsH="Acne_NL" & is.na(group_NLvsL) ~ "Increased in Acne Lesions compared to Healthy",
                                    TRUE ~ "remove")) %>% dplyr::filter(Outcome != "remove")
```

###################################################################
# Feature affected in Acne_L skin vs Acne_NL skin vs Healthy skin #
###################################################################
```{r}
feature_interest <- VIPs_LvsH_Load %>% dplyr::select(ID, comp1, GroupContrib) %>%
  dplyr::rename(vip_LvsH = "comp1", group_LvsH = "GroupContrib") %>% 
  full_join(VIPs_NLvsH_Load) %>% 
  dplyr::rename(vip_NLvsH = "comp1", group_NLvsH = "GroupContrib") %>%
  full_join(VIPs_NLvsL_Load) %>% 
  dplyr::rename(vip_NLvsL = "comp1", group_NLvsL = "GroupContrib") %>%
  dplyr::mutate(combined_vip = rowSums(cbind(vip_LvsH, vip_NLvsH, vip_NLvsL), na.rm = TRUE)) %>%
  dplyr::mutate(Group = paste(group_LvsH, group_NLvsH, group_NLvsL, sep = "_")) %>%
  dplyr::mutate(Outcome = case_when(
    # Lesional vs Healthy comparisons
    group_LvsH == "Acne_L" & is.na(group_NLvsH)  & is.na(group_NLvsL) ~ "Increased Acne_L vs Healthy",
    group_LvsH == "Healthy" & is.na(group_NLvsH)  & group_NLvsL== "Acne_NL" ~ "Increased Acne_NL and Healthy vs Acne_L",
    group_LvsH == "Healthy" & group_NLvsH == "Acne_NL"  & is.na(group_NLvsL) ~ "Increased Healthy vs Acne_L and Acne_NL vs Healthy",
    group_LvsH == "Acne_L" & is.na(group_NLvsH)  & group_NLvsL=="Acne_L" ~ "Increased Acne_L vs Healthy and Acne_NL",
    group_LvsH == "Acne_L" & group_NLvsH=="Acne_NL"  & is.na(group_NLvsL) ~ "Increased Acne_L and Acne_NL vs Healthy" ,
    group_LvsH == "Healthy" & group_NLvsH=="Healthy"  & is.na(group_NLvsL) ~ "Increased Healthy vs Acne_L and Acne_NL" ,
    group_LvsH == "Healthy" & is.na(group_NLvsH)  & is.na(group_NLvsL) ~ "Increased Healthy vs Acne_L" ,
    is.na(group_LvsH) & group_NLvsH=="Healthy"  & is.na(group_NLvsL) ~ "Increased Healthy vs Acne_NL" ,
    is.na(group_LvsH) & group_NLvsH=="Acne_NL"  & is.na(group_NLvsL) ~ "Increased Acne_NL vs Healthy" ,
    is.na(group_LvsH) & is.na(group_NLvsH)  & group_NLvsL=="Acne_L" ~ "Increased Acne_L vs Acne_NL",
    is.na(group_LvsH) & is.na(group_NLvsH)  & group_NLvsL=="Acne_NL" ~ "Increased Acne_NL vs Acne_L",
    group_LvsH == "Acne_L" & group_NLvsH=="Healthy"  & is.na(group_NLvsL) ~ "Increased Healthy vs Acne_NL and in Acne_L vs Healthy",
                                    TRUE ~ "remove")) %>% dplyr::filter(Outcome != "remove")
```


```{r}
#keep only interesting features
feature_interest <- VIPs_LvsH_Load %>% dplyr::select(ID, comp1, GroupContrib) %>%
  dplyr::rename(vip_LvsH = "comp1", group_LvsH = "GroupContrib") %>% 
  full_join(VIPs_NLvsH_Load) %>% 
  dplyr::rename(vip_NLvsH = "comp1", group_NLvsH = "GroupContrib") %>%
  full_join(VIPs_NLvsL_Load) %>% 
  dplyr::rename(vip_NLvsL = "comp1", group_NLvsL = "GroupContrib") %>%
  dplyr::mutate(combined_vip = rowSums(cbind(vip_LvsH, vip_NLvsH, vip_NLvsL), na.rm = TRUE)) %>%
  dplyr::mutate(Group = paste(group_LvsH, group_NLvsH, group_NLvsL, sep = "_")) %>%
  dplyr::mutate(Outcome = case_when(
    # Lesional vs Healthy comparisons
    group_LvsH == "Acne_L" & is.na(group_NLvsH)  & is.na(group_NLvsL) ~ "Increased Acne_L vs Healthy",
    group_LvsH == "Acne_L" & is.na(group_NLvsH)  & group_NLvsL=="Acne_L" ~ "Increased Acne_L vs Healthy and Acne_NL", #not present
    group_LvsH == "Acne_L" & group_NLvsH=="Acne_NL"  & is.na(group_NLvsL) ~ "Increased Acne_L and Acne_NL vs Healthy" ,
    group_LvsH == "Healthy" & group_NLvsH=="Healthy"  & is.na(group_NLvsL) ~ "Increased Healthy vs Acne_L and Acne_NL" ,
    is.na(group_LvsH) & is.na(group_NLvsH)  & group_NLvsL=="Acne_L" ~ "Increased Acne_L vs Acne_NL",
    group_LvsH == "Healthy" & is.na(group_NLvsH)  & group_NLvsL=="Acne_NL" ~ "Increased Healthy and Acne_NL vs Acne_L" # not present,
    
                                    TRUE ~ "remove")) %>% dplyr::filter(Outcome != "remove")
```

```{r}
summary(factor(feature_interest$Outcome))
```

```{r}
feature_Acne_Healthy <- feature_interest %>% 
  dplyr::filter(Outcome == "Increased Acne_L and Acne_NL vs Healthy")

feature_Acne_L_Healthy <- feature_interest %>% 
  dplyr::filter(Outcome == "Increased Acne_L vs Healthy")

feature_Acne_L_NL <- feature_interest %>% 
  dplyr::filter(Outcome == "Increased Acne_L vs Acne_NL")


feature_Healthy_Acne <- feature_interest %>% 
  dplyr::filter(Outcome == "Increased Healthy vs Acne_L and Acne_NL")


```

```{r}
feature_Acne <- feature_interest %>% 
  dplyr::filter(Outcome == "Increased in Acne")

feature_Healthy <- feature_interest %>% 
  dplyr::filter(Outcome == "Decreased in Acne")

feature_Acne_L_NL <- feature_interest %>% 
  dplyr::filter(Outcome == "Increased in Acne Lesions compared to Non-Lesions")


feature_Acne_L_Healthy <- feature_interest %>% 
  dplyr::filter(Outcome == "Increased in Acne Lesions compared to Healthy")

```


```{r}
# Extract exacerbated features
final_features <- feature_interest %>% 
  dplyr::filter(Outcome %in% c( "Increased in Acne", "Decreased in Acne","Increased in Acne Lesions compared to Non-Lesions", "Increased in Acne Lesions compared to Healthy"))
```


```{r}
write_csv(x = feature_interest, file = "../output/list_features_interest.csv")
write_csv(x = final_features, file = "../output/list_feaures_interest_Acne.csv")
```

